---
title: 'Polling Locations in Arizona and South Carolina: 2016 vs 2020'
author: "Jessica Randall"
date: "Last compiled `r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    toc: false
    number_sections: false
---

```{r setup, message = FALSE, include= FALSE, echo=FALSE}

# Setup ------------------------------------------------------------------------
knitr::opts_chunk$set(tidy = "styler")

pacman::p_load(
  tidyverse, knitr, janitor,
  tinytex, scales, kableExtra,
  tidycensus, stringr, sf, here
)

# Disable sci. notation in output
options(scipen = 999)

# Specify file locations -------------------------------------------------------

files <- list(
  az_2016_clean = here("write/input/az_2016_clean.rds"),
  az_2020_clean = here("write/input/az_2020_clean.rds"),
  az_demo_clean = here("write/input/az_demo_clean.rds"),
  az_places = here("write/input/az_places.rds"),

  sc_2016_clean = here("write/input/sc_2016_clean.rds"),
  sc_2020_clean = here("write/input/sc_2020_clean.rds"),
  sc_demo_clean = here("write/input/sc_demo_clean.rds"),
  sc_places = here("write/input/sc_places.rds"),

  pa_2016_clean = here("write/input/pa_2016_clean.rds"),
  pa_2020_clean = here("write/input/pa_2020_clean.rds"),
  pa_demo_clean = here("write/input/pa_demo_clean.rds"),
  pa_places = here("write/input/pa_places.rds")
)
```

## Polling locations open in 2016 and 2020 {-} 

```{r nlocationsa, message=FALSE, echo=FALSE}

# how many lost, and how many gained, maintained? ------------------------------
# AZ
az_table <- read_rds(files$az_places) %>%
  group_by(delta_cat) %>%
  summarise(
    .groups = "keep",
    count = n(),
    percent_of_zips = ((count / nrow(.)) * 100)
  )

# SC
sc_table <- read_rds(files$sc_places) %>%
  group_by(delta_cat) %>%
  summarise(
    .groups = "keep",
    count = n(),
    percent_of_zips = ((count / nrow(.)) * 100)
  )

# PA
pa_table <- read_rds(files$pa_places) %>%
  group_by(delta_cat) %>%
  summarise(
    .groups = "keep",
    count = n(),
    percent_of_zips = ((count / nrow(.)) * 100)
  )
```

```{r changes_demo, message=FALSE, echo=FALSE}

# location changes and demography ----------------------------------------------
# AZ
az <- read_rds(files$az_demo_clean)

az_lost <- az %>%
  filter(delta_cat == "Lost Polling Locations")
az_g <- az %>%
  filter(delta_cat == "Gained Polling Locations")
az_m <- az %>%
  filter(delta_cat == "Maintained Polling Locations")

az_maricopa <- az %>%
  filter(county == "Maricopa County")

az_lost_maricopa <- az_maricopa %>%
  filter(delta_cat == "Lost Polling Locations")
az_g_maricopa <- az_maricopa %>%
  filter(delta_cat == "Gained Polling Locations")
az_m_maricopa <- az_maricopa %>%
  filter(delta_cat == "Maintained Polling Locations")

# SC
sc <- read_rds(files$sc_demo_clean)

sc_lost <- sc %>%
  filter(delta_cat == "Lost Polling Locations")
sc_g <- sc %>%
  filter(delta_cat == "Gained Polling Locations")
sc_m <- sc %>%
  filter(delta_cat == "Maintained Polling Locations")

sc_rland <- sc %>%
  filter(county == "Richland")

sc_lost_rland <- sc_rland %>%
  filter(delta_cat == "Lost Polling Locations")
sc_g_rland <- sc_rland %>%
  filter(delta_cat == "Gained Polling Locations")
sc_m_rland <- sc_rland %>%
  filter(delta_cat == "Maintained Polling Locations")

ncountiessc <- sc %>%
  group_by(county) %>%
  distinct(county) %>%
  arrange(county)

# PA
pa <- read_rds(files$pa_demo_clean)

pa_lost <- pa %>%
  filter(delta_cat == "Lost Polling Locations")
pa_g <- pa %>%
  filter(delta_cat == "Gained Polling Locations")
pa_m <- pa %>%
  filter(delta_cat == "Maintained Polling Locations")

pa_inc_lost <- pa_lost %>%
  drop_na(med_income_byhh)
pa_inc_g <- pa_g %>%
  drop_na(med_income_byhh)
pa_inc_m <- pa_m %>%
  drop_na(med_income_byhh)

# Philadelphia
pa_phil <- pa %>%
  filter(county == "Philadelphia")

pa_lost_phil <- pa_phil %>%
  filter(delta_cat == "Lost Polling Locations")
pa_g_phil <- pa_phil %>%
  filter(delta_cat == "Gained Polling Locations")
pa_m_phil <- pa_phil %>%
  filter(delta_cat == "Maintained Polling Locations")

phil_inc_lost <- pa_lost_phil %>%
  drop_na(med_income_byhh)
phil_inc_g <- pa_g_phil %>%
  drop_na(med_income_byhh)
phil_inc_m <- pa_m_phil %>%
  drop_na(med_income_byhh)

# zip codes within the City of Philadelphia
city_zips <- c(
  "08102", "08103", "08105", "08110", "19004", "19012", "19027",
  "19102", "19103", "19104", "19106", "19107", "19109", "19111",
  "19119", "19120", "19121", "19122", "19123", "19124", "19125",
  "19126", "19129", "19130", "19131", "19132", "19133", "19134",
  "19135", "19137", "19138", "19140", "19141", "19144", "19147",
  "19149", "19150"
)

city_phil <- pa_phil %>%
  filter(geoid %in% city_zips)

city_lost_phil <- city_phil %>%
  filter(delta_cat == "Lost Polling Locations")
city_g_phil <- city_phil %>%
  filter(delta_cat == "Gained Polling Locations")
city_m_phil <- city_phil %>%
  filter(delta_cat == "Maintained Polling Locations")

city_inc_lost <- city_lost_phil %>%
  drop_na(med_income_byhh)
city_inc_g <- city_g_phil %>%
  drop_na(med_income_byhh)
```

```{r new_locs, message=FALSE, echo=FALSE}

# new locations in 2016 vs 2020 ------------------------------------------------

# AZ
az_2016_df <- read_rds(files$az_2016_clean)
az_2020_df <- read_rds(files$az_2020_clean)

new_addsaz <- anti_join(az_2020_df, az_2016_df, by = "address_line")

maricopa_zips <- as.character(az_maricopa$geoid)

mar_2020 <- az_2020_df %>%
  filter(zipcode %in% maricopa_zips) %>%
  distinct(address_line, .keep_all = TRUE)
# mountain vista middle azhool gym at
# 33622 north mountain vista blvd san tan valley az 85142
# is not in original maricopa 2020 data
mar_2016 <- az_2016_df %>%
  filter(zipcode %in% maricopa_zips)

new_addsaz_mar <- anti_join(mar_2020, mar_2016, by = "address_line")

# how many polling places persisted overall between 2016 and 2020?
az_both <- semi_join(az_2020_df, az_2016_df, by = "address_line")

# how many polling places persisted in maricopa between 2016 and 2020?
az_both_mar <- semi_join(mar_2020, mar_2016, by = "address_line")

# SC
sc_2016_df <- read_rds(files$sc_2016_clean)
sc_2020_df <- read_rds(files$sc_2020_clean)

new_addssc <- anti_join(sc_2020_df, sc_2016_df, by = "address_line")

richland_zips <- as.character(sc_rland$geoid)

rch_2020 <- sc_2020_df %>%
  filter(zipcode %in% richland_zips) %>%
  distinct(address_line, .keep_all = TRUE)

rch_2016 <- sc_2016_df %>%
  filter(zipcode %in% richland_zips)

# richland
new_addssc_rch <- anti_join(rch_2020, rch_2016, by = "address_line")

# how many polling places persisted overall between 2016 and 2020?
sc_both <- semi_join(sc_2020_df, sc_2016_df, by = "address_line")

# how many polling places persisted in richland between 2016 and 2020?
sc_both_rch <- semi_join(rch_2020, rch_2016, by = "address_line")

# PA
pa_2016_df <- read_rds(files$pa_2016_clean)
pa_2020_df <- read_rds(files$pa_2020_clean)
```

### Arizona

Between the 2016 primary Presidential election (PPE) and the 2020 PPE there was
a net addition of `r sum((nrow(az_2020_df) - nrow(az_2016_df)))` polling locations, 
which is a `r round_half_up((sum((nrow(az_2020_df) - nrow(az_2016_df)))/as.numeric(nrow(az_2016_df))*100),1)`% increase.

In about `r round_half_up(az_table$percent_of_zips[2])`% of the `r sum(az_table$count)`
zip codes in Arizona there was a reduction in the number of polling places. 
In total, about `r comma(sum(az_lost$total))` or 
`r round_half_up(sum(sum(az_lost$total)/sum(az$total))*100, 0)`% of the 
population of Arizona, lives in a zip code that saw a reduction in the number of 
polling places.

Compared to 2016, there were `r nrow(new_addsaz)` new polling places opened in 
Arizona statewide in 2020, `r nrow(new_addsaz_mar)` of these were opened in
Maricopa county alone. Arizona maintained `r nrow(az_both)` of the same
polling locations between 2016 and 2020 and `r nrow(az_both_mar)` of these were
within Maricopa County.

### South Carolina

In South Carolina, there was a `r abs(round_half_up((sum((nrow(sc_2020_df) - nrow(sc_2016_df)))/as.numeric(nrow(sc_2016_df))*100),1))`% reduction 
(`r abs(sum((nrow(sc_2020_df) - nrow(sc_2016_df))))`) in the number of polling 
locations between the 2016 and 2020 PPEs. In total, about 
`r comma(sum(sc_lost$total))` or 
`r round_half_up(sum(sum(sc_lost$total)/sum(sc$total))*100, 0)`% of the 
population of South Carolina, lives in a zip code that saw a reduction in the 
number of polling places.

Compared to 2016, there were `r nrow(new_addssc)` new polling places opened in 
South Carolina statewide in 2020, `r nrow(new_addssc_rch)` of these were opened in
Richland county. South Carolina maintained `r comma(nrow(sc_both))` of the same
polling locations between 2016 and 2020 and `r nrow(sc_both_rch)` of these were
within Richland County.

### Pennsylvania

Between the 2016 and 2020 PPEs there was a decrease of 
`r comma(abs(nrow(pa_2020_df) - nrow(pa_2016_df)))` polling locations, which is a
`r abs(round_half_up((sum((nrow(pa_2020_df) - nrow(pa_2016_df)))/as.numeric(nrow(pa_2016_df))*100),1))`% decrease statewide.

In about `r round_half_up(pa_table$percent_of_zips[2])`% of the 
`r comma(sum(pa_table$count))` zip codes in Pennsylvania there was a reduction 
in the number of polling places. In total, about `r comma(sum(pa_lost$total))` or 
`r round_half_up(sum(sum(pa_lost$total)/sum(pa$total))*100, 0)`% of the 
population of Pennsylvania, lives in a zip code that saw a reduction in the 
number of polling places.

## Polling location changes by demographics {-} 

### Arizona

`r round_half_up(sum(sum(az_lost$total_hl)/sum(az$total_hl)) * 100, 0)`% of
the Hispanic or Latino population lives in a zip code which lost polling locations. 
`r round_half_up(sum(sum(az_lost$nhl_white)/sum(az$nhl_white)) * 100, 0)`% of
the White population lives in a zip code which lost polling locations. 
`r round_half_up(sum(sum(az_lost$nhl_black)/sum(az$nhl_black)) * 100, 0)`% of
the Black population lives in a zip code which lost polling locations. 
`r round_half_up(sum(sum(az_lost$nhl_ai_an)/sum(az$nhl_ai_an)) * 100, 0)`% of 
the American Indian population lives in a zip code which lost polling locations.
`r round_half_up(sum(sum(az_lost$nhl_asian)/sum(az$nhl_asian)) * 100, 0)`% of 
the Asian population lives in a zip code which lost polling locations.
`r round_half_up(sum(sum(az_lost$nhl_ao)/sum(az$nhl_ao)) * 100, 0)`% 
of the population who identified as some other race or two or three or 
more races lives in a zip code which lost polling locations.

Hispanic and Latino people comprise 
`r round_half_up((sum(az$total_hl)/sum(az$total))*100,0)`% 
of the population of Arizona, and 
`r round_half_up((sum(az_lost$total_hl)/sum(az_lost$total))*100,0)`% of people 
who lived in zip codes which lost polling locations. 
White people comprise `r round_half_up((sum(az$nhl_white)/sum(az$total))*100,0)`% 
of the population of Arizona, and 
`r round_half_up((sum(az_lost$nhl_white)/sum(az_lost$total))*100,0)`% of people 
who lived in zip codes which lost polling locations. Black people comprise 
`r round_half_up((sum(az$nhl_black)/sum(az$total))*100,0)`% of the population of 
Arizona, and `r round_half_up((sum(az_lost$nhl_black)/sum(az_lost$total))*100,0)`% 
of people who lived in zip codes which lost polling locations.
American Indian people comprise 
`r round_half_up((sum(az$nhl_ai_an)/sum(az$total))*100,0)`% 
of the population of Arizona, and 
`r round_half_up((sum(az_lost$nhl_ai_an)/sum(az_lost$total))*100,0)`% of people 
who lived in zip codes which lost polling locations.
Asian people comprise 
`r round_half_up((sum(az$nhl_asian)/sum(az$total))*100,0)`% 
of the population of Arizona, and 
`r round_half_up((sum(az_lost$nhl_asian)/sum(az_lost$total))*100,0)`% of people 
who lived in zip codes which lost polling locations.

*Maricopa County*

In total, `r comma(sum(az_lost$total))`, or 
`r round_half_up(sum(sum(az_lost_maricopa$total)/sum(az_maricopa$total))*100,0)`% 
of people who lived in Maricopa County, Arizona lived in zip codes which lost 
polling locations during the 2020 PPE.

`r round_half_up(sum(sum(az_lost_maricopa$total_hl)/sum(az_maricopa$total_hl)) * 100, 0)`% of
the Hispanic or Latino population lives in a zip code which lost polling locations. 
`r round_half_up(sum(sum(az_lost_maricopa$nhl_white)/sum(az_maricopa$nhl_white)) * 100, 0)`% of
the White population lives in a zip code which lost polling locations. 
`r round_half_up(sum(sum(az_lost_maricopa$nhl_black)/sum(az_maricopa$nhl_black)) * 100, 0)`% of
the Black population lives in a zip code which lost polling locations. 
`r round_half_up(sum(sum(az_lost_maricopa$nhl_ai_an)/sum(az_maricopa$nhl_ai_an)) * 100, 0)`% of 
the American Indian population lives in a zip code which lost polling locations.
`r round_half_up(sum(sum(az_lost_maricopa$nhl_asian)/sum(az_maricopa$nhl_asian)) * 100, 0)`% of 
the Asian population lives in a zip code which lost polling locations.
`r round_half_up(sum(sum(az_lost_maricopa$nhl_ao)/sum(az_maricopa$nhl_ao)) * 100, 0)`% 
of the population who identified as some other race or two or three or 
more races lives in a zip code which lost polling locations.

Hispanic and Latino people comprise 
`r round_half_up((sum(az_maricopa$total_hl)/sum(az_maricopa$total))*100,0)`% 
of the population of Maricopa County, and 
`r round_half_up((sum(az_lost_maricopa$total_hl)/sum(az_lost_maricopa$total))*100,0)`% 
of people who lived in zip codes which lost polling locations. 
White people comprise `r round_half_up((sum(az_maricopa$nhl_white)/sum(az_maricopa$total))*100,0)`% 
of the population of Maricopa County, and 
`r round_half_up((sum(az_lost_maricopa$nhl_white)/sum(az_lost_maricopa$total))*100,0)`% 
of people who lived in zip codes which lost polling locations. Black people 
comprise `r round_half_up((sum(az_maricopa$nhl_black)/sum(az_maricopa$total))*100,0)`% of 
the population of Maricopa County, and `r round_half_up((sum(az_lost_maricopa$nhl_black)/sum(az_lost_maricopa$total))*100,0)`% 
of people who lived in zip codes which lost polling locations. American Indian 
people comprise `r round_half_up((sum(az_maricopa$nhl_ai_an)/sum(az_maricopa$total))*100,0)`% 
of the population of Maricopa County, and 
`r round_half_up((sum(az_lost_maricopa$nhl_ai_an)/sum(az_lost_maricopa$total))*100,0)`% 
of people who lived in zip codes which lost polling locations. Asian people 
comprise `r round_half_up((sum(az_maricopa$nhl_asian)/sum(az_maricopa$total))*100,0)`% 
of the population of Maricopa County, and 
`r round_half_up((sum(az_lost_maricopa$nhl_asian)/sum(az_lost_maricopa$total))*100,0)`% 
of people who lived in zip codes which lost polling locations.

### South Carolina

`r round_half_up(sum(sum(sc_lost$total_hl)/sum(sc$total_hl)) * 100, 0)`% of
the Hispanic or Latino population lives in a zip code which lost polling locations. 
`r round_half_up(sum(sum(sc_lost$nhl_white)/sum(sc$nhl_white)) * 100, 0)`% of
the White population lives in a zip code which lost polling locations. 
`r round_half_up(sum(sum(sc_lost$nhl_black)/sum(sc$nhl_black)) * 100, 0)`% of
the Black population lives in a zip code which lost polling locations. 
`r round_half_up(sum(sum(sc_lost$nhl_ai_an)/sum(sc$nhl_ai_an)) * 100, 0)`% of 
the American Indian population lives in a zip code which lost polling locations.
`r round_half_up(sum(sum(sc_lost$nhl_asian)/sum(sc$nhl_asian)) * 100, 0)`% of 
the Asian population lives in a zip code which lost polling locations.
`r round_half_up(sum(sum(sc_lost$nhl_ao)/sum(sc$nhl_ao)) * 100, 0)`% 
of the population who identified as some other race or two or three or 
more races lives in a zip code which lost polling locations.

Hispanic and Latino people comprise 
`r round_half_up((sum(sc$total_hl)/sum(sc$total))*100,0)`% 
of the population of South Carolina, and 
`r round_half_up((sum(sc_lost$total_hl)/sum(sc_lost$total))*100,0)`% of people 
who lived in zip codes which lost polling locations. 
White people comprise `r round_half_up((sum(sc$nhl_white)/sum(sc$total))*100,0)`% 
of the population of South Carolina, and 
`r round_half_up((sum(sc_lost$nhl_white)/sum(sc_lost$total))*100,0)`% of people 
who lived in zip codes which lost polling locations. Black people comprise 
`r round_half_up((sum(sc$nhl_black)/sum(sc$total))*100,0)`% of the population of 
South Carolina, and `r round_half_up((sum(sc_lost$nhl_black)/sum(sc_lost$total))*100,0)`% 
of people who lived in zip codes which lost polling locations.
American Indian people comprise less than 1% of the population of South Carolina, 
and fewer than 1 % of people who lived in zip codes which lost polling locations.
Asian people comprise 
`r round_half_up((sum(sc$nhl_asian)/sum(sc$total))*100,0)`% 
of the population of South Carolina, and 
`r round_half_up((sum(sc_lost$nhl_asian)/sum(sc_lost$total))*100,0)`% of people 
who lived in zip codes which lost polling locations.

*Richland County*

In total, `r comma(sum(sc_lost$total))`, or 
`r round_half_up(sum(sum(sc_lost_rland$total)/sum(sc_rland$total))*100,0)`% of 
people who lived in Richland County, South Carolina lived in zip codes which 
lost polling locations during the 2020 PPE.

`r round_half_up(sum(sum(sc_lost_rland$total_hl)/sum(sc_rland$total_hl)) * 100, 0)`% of
the Hispanic or Latino population lives in a zip code which lost polling locations. 
`r round_half_up(sum(sum(sc_lost_rland$nhl_white)/sum(sc_rland$nhl_white)) * 100, 0)`% of
the White population lives in a zip code which lost polling locations. 
`r round_half_up(sum(sum(sc_lost_rland$nhl_black)/sum(sc_rland$nhl_black)) * 100, 0)`% of
the Black population lives in a zip code which lost polling locations. 
`r round_half_up(sum(sum(sc_lost_rland$nhl_ai_an)/sum(sc_rland$nhl_ai_an)) * 100, 0)`% of 
the American Indian population lives in a zip code which lost polling locations.
`r round_half_up(sum(sum(sc_lost_rland$nhl_asian)/sum(sc_rland$nhl_asian)) * 100, 0)`% of 
the Asian population lives in a zip code which lost polling locations.
`r round_half_up(sum(sum(sc_lost_rland$nhl_ao)/sum(sc_rland$nhl_ao)) * 100, 0)`% 
of the population who identified as some other race or two or three or 
more races lives in a zip code which lost polling locations.

Hispanic and Latino people comprise 
`r round_half_up((sum(sc_rland$total_hl)/sum(sc_rland$total))*100,0)`% 
of the population of Richland County, and 
`r round_half_up((sum(sc_lost_rland$total_hl)/sum(sc_lost_rland$total))*100,0)`% 
of people who lived in zip codes which lost polling locations. 
White people comprise `r round_half_up((sum(sc_rland$nhl_white)/sum(sc_rland$total))*100,0)`% 
of the population of Richland County, and 
`r round_half_up((sum(sc_lost_rland$nhl_white)/sum(sc_lost_rland$total))*100,0)`% 
of people who lived in zip codes which lost polling locations. Black people 
comprise `r round_half_up((sum(sc_rland$nhl_black)/sum(sc_rland$total))*100,0)`% of 
the population of Richland County, and `r round_half_up((sum(sc_lost_rland$nhl_black)/sum(sc_lost_rland$total))*100,0)`% 
of people who lived in zip codes which lost polling locations. American Indian 
people comprise less than 1% of the population of Richland County and fewer 
than 1 % of people who lived in zip codes which lost polling locations. Asian 
people comprise 
`r round_half_up((sum(sc_rland$nhl_asian)/sum(sc_rland$total))*100,0)`% 
of the population of Richland County, and 
`r round_half_up((sum(sc_lost_rland$nhl_asian)/sum(sc_lost_rland$total))*100,0)`% 
of people who lived in zip codes which lost polling locations.

### Pennsylvania

`r round_half_up(sum(sum(pa_lost$total_hl)/sum(pa$total_hl)) * 100, 0)`% of
the Hispanic or Latino population live in a zip code which lost polling locations. 
`r round_half_up(sum(sum(pa_lost$nhl_white)/sum(pa$nhl_white)) * 100, 0)`% of
the White population live in a zip code which lost polling locations. 
`r round_half_up(sum(sum(pa_lost$nhl_black)/sum(pa$nhl_black)) * 100, 0)`% of
the Black population live in a zip code which lost polling locations. 
`r round_half_up(sum(sum(pa_lost$nhl_ai_an)/sum(pa$nhl_ai_an)) * 100, 0)`% of 
the American Indian population live in a zip code which lost polling locations.
`r round_half_up(sum(sum(pa_lost$nhl_asian)/sum(pa$nhl_asian)) * 100, 0)`% of 
the Asian population live in a zip code which lost polling locations.
`r round_half_up(sum(sum(pa_lost$nhl_ao)/sum(pa$nhl_ao)) * 100, 0)`% 
of the population who identified as some other race or two or three or 
more races live in a zip code which lost polling locations.

Hispanic and Latino people comprise 
`r round_half_up((sum(pa$total_hl)/sum(pa$total))*100,0)`% 
of the population of Pennsylvania, and 
`r round_half_up((sum(pa_lost$total_hl)/sum(pa_lost$total))*100,0)`% of people 
who lived in zip codes which lost polling locations. 
White people comprise `r round_half_up((sum(pa$nhl_white)/sum(pa$total))*100,0)`% 
of the population of Pennsylvania, and 
`r round_half_up((sum(pa_lost$nhl_white)/sum(pa_lost$total))*100,0)`% of people 
who lived in zip codes which lost polling locations. Black people comprise 
`r round_half_up((sum(pa$nhl_black)/sum(pa$total))*100,0)`% of the population of 
Pennsylvania, and `r round_half_up((sum(pa_lost$nhl_black)/sum(pa_lost$total))*100,0)`% 
of people who lived in zip codes which lost polling locations.
American Indian people comprise less than 1% of the population of Pennsylvania, 
and fewer than 1% of people who lived in zip codes which lost polling locations.
Asian people comprise 
`r round_half_up((sum(pa$nhl_asian)/sum(pa$total))*100,0)`% 
of the population of Pennsylvania, and 
`r round_half_up((sum(pa_lost$nhl_asian)/sum(pa_lost$total))*100,0)`% of people 
who lived in zip codes which lost polling locations.

The weighted median income in the previous year of all zip codes which lost 
polling places between 2016 and 2020 was 
\$`r comma(median(pa_inc_lost$med_income_byhh))` and the weighted median income 
of all zip codes in Pennsylvania which gained polling locations was 
\$`r comma(median(pa_inc_g$med_income_byhh))` and among those which
maintained polling locations it was \$`r comma(median(pa_inc_m$med_income_byhh))`.

*Philadelphia County*

In total, `r comma(sum(pa_lost$total))`, or 
`r round_half_up(sum(sum(pa_lost_phil$total)/sum(pa_phil$total))*100,0)`% of 
people who lived in Philadelphia County, Pennsylvania lived in zip codes which 
lost polling locations during the 2020 PPE.

`r round_half_up(sum(sum(pa_lost_phil$total_hl)/sum(pa_phil$total_hl)) * 100, 0)`% of
the Hispanic or Latino population lives in a zip code which lost polling locations. 
`r round_half_up(sum(sum(pa_lost_phil$nhl_white)/sum(pa_phil$nhl_white)) * 100, 0)`% of
the White population lives in a zip code which lost polling locations. 
`r round_half_up(sum(sum(pa_lost_phil$nhl_black)/sum(pa_phil$nhl_black)) * 100, 0)`% of
the Black population lives in a zip code which lost polling locations. 
`r round_half_up(sum(sum(pa_lost_phil$nhl_ai_an)/sum(pa_phil$nhl_ai_an)) * 100, 0)`% of 
the American Indian population lives in a zip code which lost polling locations.
`r round_half_up(sum(sum(pa_lost_phil$nhl_asian)/sum(pa_phil$nhl_asian)) * 100, 0)`% of 
the Asian population lives in a zip code which lost polling locations.
`r round_half_up(sum(sum(pa_lost_phil$nhl_ao)/sum(pa_phil$nhl_ao)) * 100, 0)`% 
of the population who identified as some other race or two or three or 
more races lives in a zip code which lost polling locations.

Hispanic and Latino people comprise 
`r round_half_up((sum(pa_phil$total_hl)/sum(pa_phil$total))*100,0)`% 
of the population of Philadelphia County, and 
`r round_half_up((sum(pa_lost_phil$total_hl)/sum(pa_lost_phil$total))*100,0)`% 
of people who lived in zip codes which lost polling locations. 
White people comprise `r round_half_up((sum(pa_phil$nhl_white)/sum(pa_phil$total))*100,0)`% 
of the population of Philadelphia County, and 
`r round_half_up((sum(pa_lost_phil$nhl_white)/sum(pa_lost_phil$total))*100,0)`% 
of people who lived in zip codes which lost polling locations. Black people 
comprise `r round_half_up((sum(pa_phil$nhl_black)/sum(pa_phil$total))*100,0)`% 
of the population of Philadelphia County, and `r round_half_up((sum(pa_lost_phil$nhl_black)/sum(pa_lost_phil$total))*100,0)`% 
of people who lived in zip codes which lost polling locations. American Indian 
people comprise less than 1% of the population of Philadelphia County, and fewer 
than 1 % of people who lived in zip codes which lost polling locations. 
Asian people comprise 
`r round_half_up((sum(pa_phil$nhl_asian)/sum(pa_phil$total))*100,0)`% 
of the population of Philadelphia County, and 
`r round_half_up((sum(pa_lost_phil$nhl_asian)/sum(pa_lost_phil$total))*100,0)`% 
of people who lived in zip codes which lost polling locations.

The weighted median income in the previous year of all zip codes which lost 
polling places between 2016 and 2020 was 
\$`r comma(median(phil_inc_lost$med_income_byhh))` and the weighted median 
income of all zip codes in Philadelphia County which gained polling locations 
was \$`r comma(median(phil_inc_g$med_income_byhh))` and among those which
maintained polling locations it was \$`r comma(median(phil_inc_m$med_income_byhh))`.

*City of Philadelphia*

In total, `r comma(sum(city_lost_phil$total))`, or 
`r round_half_up(sum(sum(city_lost_phil$total)/sum(pa_phil$total))*100,0)`% of 
people who lived within the City of Philadelphia lived in zip codes which lost 
polling locations during the 2020 PPE.

`r round_half_up(sum(sum(city_lost_phil$total_hl)/sum(pa_phil$total_hl)) * 100, 0)`% of
the Hispanic or Latino population lives in a zip code which lost polling locations. 
`r round_half_up(sum(sum(city_lost_phil$nhl_white)/sum(pa_phil$nhl_white)) * 100, 0)`% of
the White population lives in a zip code which lost polling locations. 
`r round_half_up(sum(sum(city_lost_phil$nhl_black)/sum(pa_phil$nhl_black)) * 100, 0)`% of
the Black population lives in a zip code which lost polling locations. 
`r round_half_up(sum(sum(city_lost_phil$nhl_ai_an)/sum(pa_phil$nhl_ai_an)) * 100, 0)`% of 
the American Indian population lives in a zip code which lost polling locations.
`r round_half_up(sum(sum(city_lost_phil$nhl_asian)/sum(pa_phil$nhl_asian)) * 100, 0)`% of 
the Asian population lives in a zip code which lost polling locations.
`r round_half_up(sum(sum(city_lost_phil$nhl_ao)/sum(pa_phil$nhl_ao)) * 100, 0)`% 
of the population who identified as some other race or two or three or 
more races lives in a zip code which lost polling locations.

Hispanic and Latino people comprise 
`r round_half_up((sum(city_phil$total_hl)/sum(city_phil$total))*100,0)`% 
of the population of City of Philadelphia, and 
`r round_half_up((sum(pa_lost_phil$total_hl)/sum(pa_lost_phil$total))*100,0)`% 
of people who lived in zip codes which lost polling locations. 
White people comprise `r round_half_up((sum(city_phil$nhl_white)/sum(city_phil$total))*100,0)`% 
of the population of City of Philadelphia, and 
`r round_half_up((sum(pa_lost_phil$nhl_white)/sum(pa_lost_phil$total))*100,0)`% 
of people who lived in zip codes which lost polling locations. Black people 
comprise `r round_half_up((sum(city_phil$nhl_black)/sum(city_phil$total))*100,0)`% 
of the population of City of Philadelphia, and `r round_half_up((sum(pa_lost_phil$nhl_black)/sum(pa_lost_phil$total))*100,0)`% 
of people who lived in zip codes which lost polling locations. American Indian 
people comprise less than 1% of the population of City of Philadelphia, and 
fewer than 1 % of people who lived in zip codes which lost polling locations. 
Asian people comprise 
`r round_half_up((sum(city_phil$nhl_asian)/sum(city_phil$total))*100,0)`% 
of the population of City of Philadelphia, and 
`r round_half_up((sum(pa_lost_phil$nhl_asian)/sum(pa_lost_phil$total))*100,0)`% 
of people who lived in zip codes which lost polling locations.

The weighted median income in the previous year of all zip codes which lost 
polling places between 2016 and 2020 was 
\$`r comma(median(city_inc_lost$med_income_byhh))` and the weighted median 
income of all zip codes in the City of Philadelphia which gained polling 
locations was \$`r comma(median(city_inc_g$med_income_byhh))` and there were no 
zip codes within the City of Philadelphia which maintained the same number of 
polling locations between 2016 and 2020.

## Polling location changes by population {-} 

*Selected Tables*

## Arizona

The following tables show the number of people compared with how many zip codes 
gained, maintained, or lost polling locations.In zip codes where between 
0 -- 10,000 people live, 52 gained polling locations, 10 lost polling locations, 
and 84 maintained polling locations.

```{r azkable, echo=FALSE, message=FALSE}

az_t1 <- az %>%
  tabyl(gp_total, delta_cat) %>%
  adorn_totals("row") %>%
  adorn_title("combined")

kable(az_t1, "latex",
  booktabs = T, row.names = TRUE,
  caption = "Arizona"
) %>%
  kable_styling(position = "center", latex_options = "HOLD_position")
```

Zip codes with between 0 -- 10,000 people gained the greatest number of polling 
locations between the 2016 and 2020 PPE. Those with between 30,001 -- 40,000 
people gained the second highest number of polling locations, and zip codes with 
between 70,001 -- 80,000 people gained the fewest polling locations.

Zip codes with between 0 -- 10,000 people maintained the greatest number of 
polling locations between the 2016 and 2020 PPE, those with between 
10,001 -- 20,000 people maintained the second highest number of polling 
locations, and zip codes with between 70,001 -- 80,000 people maintained no 
polling locations. 

Zip codes with between 0 -- 10,000 people lost the greatest number of polling 
locations between the 2016 and 2020 PPE, those with between 10,001 -- 20,000 
people lost the second highest number of polling locations, and zip codes with 
between 50,001 -- 60,000 and between 70,001 -- 80,000 people lost no
polling locations.

## South Carolina 

```{r sckable, echo=FALSE, message=FALSE}

sc_t1 <- sc %>%
  tabyl(gp_total, delta_cat) %>%
  adorn_totals("row") %>%
  adorn_title("combined")

kable(sc_t1, "latex",
  booktabs = T, row.names = TRUE,
  caption = "South Carolina"
) %>%
  kable_styling(position = "center", latex_options = "HOLD_position")
```

Zip codes with between 0 -- 10,000 people and those with between 
10,001 -- 20,000 people gained the greatest number of polling locations between 
the 2016 and 2020 PPE. Those with between 20,001 -- 30,000 people gained the 
second highest number of polling locations, and zip codes with between 
60,001 -- 70,000 people and those with between 80,001 -- 90,000 people
gained no polling locations.

Zip codes with between 0 -- 10,000 people maintained the greatest number of 
polling locations between the 2016 and 2020 PPE, those with between 
10,001 -- 20,000 people maintained the second highest number of polling 
locations, and zip codes with between 80,001 -- 90,000 people maintained no 
locations. 

Zip codes with between 10,001 -- 20,000 people lost the greatest number of 
polling locations between the 2016 and 2020 PPE, those with between 0 -- 10,000 
people lost the second highest number of polling locations, and zip codes with 
between 60,001 -- 70,000 people and those with between 80,001 -- 90,000 people
lost the fewest locations. 

## Pennsylvania

```{r pakable, echo=FALSE, message=FALSE}

pa_t1 <- pa %>%
  tabyl(gp_total, delta_cat) %>%
  adorn_totals("row") %>%
  adorn_title("combined")

kable(pa_t1, "latex",
  booktabs = T, row.names = TRUE,
  caption = "Pennsylvania"
) %>%
  kable_styling(position = "center", latex_options = "HOLD_position")
```

Zip codes with between 0 -- 10,000 people gained the greatest number of polling 
locations between the 2016 and 2020 PPE. Those with between 
10,001 -- 20,000 people gained the second highest number of polling locations, 
and zip codes with between 70,001 -- 80,000 people gained the fewest number of
locations.

Zip codes with between 0 -- 10,000 people maintained the greatest number of 
polling locations between the 2016 and 2020 PPE, those with between 
10,001 -- 20,000 people maintained the second highest number of polling locations, 
and zip codes with between both 60,001 -- 70,000 people and those with between 
70,001 -- 80,000 people maintained no locations. 

Zip codes with between 10,001 -- 20,000 people lost the greatest number of 
polling locations between the 2016 and 2020 PPE, those with between 
10,001 -- 20,000 people lost the second highest number of polling locations, and 
zip codes with between both 60,001 -- 70,000 people and those with between 
70,001 -- 80,000 people lost no locations. 

## Methodology

``` {r ncounties, echo=FALSE, message=FALSE}

# for how many counties do we have data? ---------------------------------------

ncountiesaz <- az %>%
  group_by(county) %>%
  distinct(county) %>%
  arrange(county)

ncountiessc <- sc %>%
  group_by(county) %>%
  distinct(county) %>%
  arrange(county)

ncountiespa <- pa %>%
  group_by(county) %>%
  distinct(county) %>%
  arrange(county)
```

This report examines polling location data from `r nrow(ncountiesaz)`
or `r round_half_up(sum((nrow(ncountiesaz)/15)*100),0)`% of the counties in 
Arizona, `r nrow(ncountiessc)` or `r round_half_up(sum((nrow(ncountiessc)/46)*100),0)`% 
of the counties in South Carolina, and `r nrow(ncountiespa)` or `r round_half_up(sum((nrow(ncountiespa)/67)*100),0)`% of the counties in Pennsylvania.

Pima County, Arizona has not been included since the data come from the Voter 
Information Project and Pima County does not participate in its activities. 
Data were reviewed for uniqueness, completeness, and accuracy of addresses.
Polling locations are known to change just before an election so this data 
should not be viewed as a definitive list of the actual polling locations used 
on election day.

Demographic data were obtained from the 2014 -- 2018 American Community Survey's 
B03002 table and groups were categorized as either Hispanic or Latino, and
not Hispanic or Latino by specific group: White alone, Black or African American 
Alone, American Indian and Alaska Native alone, Asian alone, Native Hawaiian and 
Other Pacific Islander alone, and All other races alone (Including Some other 
race alone and Two or more races alone).

Weighted median household income calculations were estimated by taking the 
median of the median household incomes by household size as estimated in the 
American Community Survey table B19019_001. Only zip codes with complete 
information on median income by household size were included in income 
calculations so these numbers may differ from total number of zip codes which 
saw changes in numbers of polling locations.

Zip code and county data were obtained from unitedstateszipcodes.org. 
While many zip codes map directly to specific counties, the relationship between 
these two organizational systems is not always 1:1 and is subject to change.
Data are current as of as of `r format(Sys.time(), '%d %B, %Y')`.

## Acknowledgements {-} 

We are grateful to Sarah Berlin at Democracy Works for providing data on Arizona 
and South Carolina from the Voter Information Project and Rebecca Fuhrman at the 
Commonwealth of Pennsylvania's Department of State for providing data from 
Pennsylvania.
