---
title: 'Polling Locations in Arizona and South Carolina: 2016 vs 2020'
author: "Jessica Randall"
date: "Last compiled `r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    toc: false
    number_sections: false
---

```{r setup, message = FALSE, include= FALSE, echo=FALSE}

pacman::p_load(tidyverse, knitr, janitor, tinytex, scales)
# remove scientific notation from output
options(scipen = 999)

files <- list(
  az_2016_clean = here::here("write/input/az_2016_clean.rds"),
  az_2016_freq_clean = here::here("write/input/az_2016_freq_clean.rds"),
  az_2020_clean = here::here("write/input/az_2020_clean.rds"),
  az_2020_freq_clean = here::here("write/input/az_2020_freq_clean.rds"),
  az_demo_clean = here::here("write/input/az_demo_clean.rds"),
  az_demo_covid_clean = here::here("write/input/az_demo_covid_clean.rds"),
  sc_2016_clean = here::here("write/input/sc_2016_clean.rds"),
  sc_2016_freq_clean = here::here("write/input/sc_2016_freq_clean.rds"),
  sc_2020_clean = here::here("write/input/sc_2020_clean.rds"),
  sc_2020_freq_clean = here::here("write/input/sc_2020_freq_clean.rds"),
  sc_demo_clean = here::here("write/input/sc_demo_clean.rds"), 
  sc_demo_covid_clean = here::here("write/input/sc_demo_covid_clean.rds")
)

```

## Polling locations open in 2016 and 2020 {-} 

```{r nlocationsa, message=FALSE, echo=FALSE}

az_2016_df <- read_rds(files$az_2016_clean)
az_2020_df <- read_rds(files$az_2020_clean)

sc_2016_df <- read_rds(files$sc_2016_clean)
sc_2020_df <- read_rds(files$sc_2020_clean)

```

### Arizona

Overall, Arizona had `r nrow(az_2016_df)` open polling locations statewide at the 2016
primary Presidential election (PPE). During the 2020 PPE there were `r nrow(az_2020_df)` 
open polling locations. This is a difference of `r sum((nrow(az_2020_df) - nrow(az_2016_df)))` 
more or a `r round_half_up((sum((nrow(az_2020_df) - nrow(az_2016_df)))/as.numeric(nrow(az_2016_df))*100),1)`% 
increase in polling locations.

### South Carolina

Overall, South Carolina had `r nrow(sc_2016_df)` open polling locations statewide at the 2016
PPE. During the 2020 PPE there were `r nrow(sc_2020_df)` open polling locations. 
This is a difference of `r abs(sum((nrow(sc_2020_df) - nrow(sc_2016_df))))` fewer 
or a `r abs(round_half_up((sum((nrow(sc_2020_df) - nrow(sc_2016_df)))/as.numeric(nrow(sc_2016_df))*100),1))`% 
decrease in polling locations.

## Changes polling locations between 2016 and 2020 {-} 

```{r nlocationsb, message=FALSE, echo=FALSE}
# AZ
az_zips_freq_2016 <- read_rds(files$az_2016_freq_clean)
az_zips_freq_2020 <- read_rds(files$az_2020_freq_clean)

n_places_az <- full_join(az_zips_freq_2020, az_zips_freq_2016, by = "zipcode") %>%
  replace_na(list(n_pp_2016 = 0, n_pp_2020 = 0)) %>%
  mutate(delta_n_places = as.numeric(n_pp_2020 - n_pp_2016),
         delta_cat = as.factor(if_else(delta_n_places < 0, 
                                       "Lost Polling Places", 
                                       "Gained or Maintained Polling Places"))) %>%
  arrange(delta_n_places) %>%
  select(zipcode, n_pp_2020, n_pp_2016, delta_n_places, delta_cat)

# how many lost, and how many gained, maintained?
az_table <- n_places_az %>%
  group_by(delta_cat) %>%
  summarise(.groups = "keep", 
            count = n(),
            percent_of_zips = ((count/nrow(n_places_az))*100))

# SC
sc_zips_freq_2016 <- read_rds(files$sc_2016_freq_clean)
sc_zips_freq_2020 <- read_rds(files$sc_2020_freq_clean)

# which zip codes saw an increase, decrease, or maintenance in polling places?
n_places_sc <- full_join(sc_zips_freq_2020, sc_zips_freq_2016, by = "zipcode") %>%
  replace_na(list(n_pp_2016 = 0, n_pp_2020 = 0)) %>%
  mutate(delta_n_places = as.numeric(n_pp_2020 - n_pp_2016),
         delta_cat = as.factor(if_else(delta_n_places < 0, 
                                       "Lost Polling Places", 
                                       "Gained or Maintained Polling Places"))) %>%
  filter(zipcode != "SC") %>%
  arrange(delta_n_places) %>%
  select(zipcode, n_pp_2020, n_pp_2016, delta_n_places, delta_cat)

# how many lost, and how many gained, maintained?
sc_table <- n_places_sc %>%
  group_by(delta_cat) %>%
  summarise(.groups = "keep", 
            count = n(),
            percent_of_zips = (count/nrow(n_places_sc)*100))

```

### Arizona 

Between the 2016 and 2020 PPEs, `r az_table$count[1]` zip codes, or 
`r round_half_up(az_table$percent_of_zips[1],1)`% of zip codes in Arizona with 
open polling locations during the 2016 PPE gained or maintained the same number 
of polling locations during the 2020 PPE while `r az_table$count[2]` zip codes, 
or `r round_half_up(az_table$percent_of_zips[2],1)`%, lost polling locations 
during the 2020 PPE.

### South Carolina 
  
Between the 2016 and 2020 PPEs, `r sc_table$count[1]` zip codes, or 
`r round_half_up(sc_table$percent_of_zips[1],1)`% of zip codes in South Carolina 
with open polling locations during the 2016 PPE gained or maintained the same 
number of polling locations during the 2020 PPE while `r sc_table$count[2]` zip 
codes, or `r round_half_up(sc_table$percent_of_zips[2],1)`%, lost polling 
locations during the 2020 PPE.
 
## Polling location changes and demographics {-} 

```{r demographics, message=FALSE, echo=FALSE}

az_demo <- read_rds(files$az_demo_clean)
az_demo_lost <- az_demo %>%
  filter(delta_cat == "Lost Polling Places")
az_l_maricopa <- az_demo_lost %>%
  filter(county == "Maricopa County")
az_demo_gm <- az_demo %>%
  filter(delta_cat == "Gained or Maintained Polling Places")
az_l_maricopa <- az_demo_lost %>%
  filter(county == "Maricopa County")
az_gm_maricopa <- az_demo_gm %>%
  filter(county == "Maricopa County")

ncountiesaz <- az_demo %>%
  group_by(county) %>%
  distinct(county) %>%
  arrange(county)

sc_demo <- read_rds(files$sc_demo_clean)
sc_demo_lost <- sc_demo %>%
  filter(delta_cat == "Lost Polling Places")
sc_demo_gm <- sc_demo %>%
  filter(delta_cat == "Gained or Maintained Polling Places")
sc_l_rland <- sc_demo_lost %>%
  filter(county == "Richland")
sc_gm_rland <- sc_demo_gm %>%
  filter(county == "Richland")

ncountiessc <- sc_demo %>%
  group_by(county) %>%
  distinct(county) %>%
  arrange(county)

```

### Arizona

In those zip codes which lost polling locations (n = `r nrow(az_demo_lost)`), 
an average of `r round_half_up(mean(az_demo_lost$pct_hl),0)`% (median = 
`r round_half_up(median(az_demo_lost$pct_hl),0)`%) of people considered 
themselves Hispanic or Latino, an average of `r round_half_up(mean(az_demo_lost$pct_nhl_white),0)`% 
(median = `r round_half_up(median(az_demo_lost$pct_nhl_white),0)`%) of people 
considered themselves White, an average of `r round_half_up(mean(az_demo_lost$pct_nhl_black),0)`% 
(median = `r round_half_up(median(az_demo_lost$pct_nhl_black),0)`%) of people 
considered themselves Black or African-American, an average of `r round_half_up(mean(az_demo_lost$pct_nhl_ai_an),0)`% 
(median = `r round_half_up(median(az_demo_lost$pct_nhl_ai_an),0)`%) of people 
considered themselves American Indian or Alaskan Native, an average of `r round_half_up(mean(az_demo_lost$pct_nhl_asian),0)`% 
(median = `r round_half_up(median(az_demo_lost$pct_nhl_asian),0)`%) of people 
considered themselves Asian, fewer than 1% of of people considered themselves 
Native Hawaiian or Pacific Islander, and an average of `r round_half_up(mean(az_demo_lost$pct_nhl_ao),0)`% 
(median = `r round_half_up(median(az_demo_lost$pct_nhl_ao),0)`%) of people 
considered themselves some other race or two or three or more races.

In those zip codes which gained or maintained polling locations (n = `r nrow(az_demo_gm)`), 
an average of `r round_half_up(mean(az_demo_gm$pct_hl),0)`% (median = `r round_half_up(median(az_demo_gm$pct_hl),0)`%) 
of people considered themselves Hispanic or Latino, an average of `r round_half_up(mean(az_demo_gm$pct_nhl_white),0)`% 
(median = `r round_half_up(median(az_demo_gm$pct_nhl_white),0)`%) of people 
considered themselves White, an average of `r round_half_up(mean(az_demo_gm$pct_nhl_black),0)`% 
(median = `r round_half_up(median(az_demo_gm$pct_nhl_black),0)`%) of people 
considered themselves Black or African-American, an average of `r round_half_up(mean(az_demo_gm$pct_nhl_ai_an),0)`% 
(median = `r round_half_up(median(az_demo_gm$pct_nhl_ai_an),0)`%) of people 
considered themselves American Indian or Alaskan Native, an average of `r round_half_up(mean(az_demo_gm$pct_nhl_asian),0)`% 
(median = `r round_half_up(median(az_demo_gm$pct_nhl_asian),0)`%) of people 
considered themselves Asian, fewer than 1% of people considered themselves 
Native Hawaiian or Pacific Islander, and an average of `r round_half_up(mean(az_demo_gm$pct_nhl_ao),0)`% 
(median = `r round_half_up(median(az_demo_gm$pct_nhl_ao),0)`%) of people 
considered themselves some other race or two or three or more races.

*Maricopa County*

In those zip codes which lost polling locations within Maricopa County 
(n = `r nrow(az_l_maricopa)`), an average of `r round_half_up(mean(az_l_maricopa$pct_hl),0)`% 
(median = `r round_half_up(median(az_l_maricopa$pct_hl),0)`%) of people 
considered themselves Hispanic or Latino, an average of `r round_half_up(mean(az_l_maricopa$pct_nhl_white),0)`% (median = 
`r round_half_up(median(az_l_maricopa$pct_nhl_white),0)`%) of people considered 
themselves White, an average of `r round_half_up(mean(az_l_maricopa$pct_nhl_black),0)`% (median = 
`r round_half_up(median(az_l_maricopa$pct_nhl_black),0)`%) of people considered
themselves Black or African-American, an average of `r round_half_up(mean(az_l_maricopa$pct_nhl_ai_an),0)`% 
(median = `r round_half_up(median(az_l_maricopa$pct_nhl_ai_an),0)`%) of people considered
themselves American Indian or Alaskan Native, an average of `r round_half_up(mean(az_l_maricopa$pct_nhl_asian),0)`% 
(median = `r round_half_up(median(az_l_maricopa$pct_nhl_asian),0)`%) of people 
considered themselves Asian, fewer than 1% of people considered themselves 
Native Hawaiian or Pacific Islander, and an average of `r round_half_up(mean(az_l_maricopa$pct_nhl_ao),0)`% 
(median = `r round_half_up(median(az_l_maricopa$pct_nhl_ao),0)`%) of people 
considered themselves some other race or two or three or more races.

In those zip codes which gained or maintained polling locations within 
Maricopa County (n = `r nrow(az_gm_maricopa)`), an average of `r round_half_up(mean(az_gm_maricopa$pct_hl),0)`% 
(median = `r round_half_up(median(az_gm_maricopa$pct_hl),0)`%) of people 
considered themselves Hispanic or Latino, an average of `r round_half_up(mean(az_gm_maricopa$pct_nhl_white),0)`% 
(median = `r round_half_up(median(az_gm_maricopa$pct_nhl_white),0)`%) of people 
considered themselves White, an average of `r round_half_up(mean(az_gm_maricopa$pct_nhl_black),0)`% 
(median = `r round_half_up(median(az_gm_maricopa$pct_nhl_black),0)`%) of people 
considered themselves Black or African-American, an average of  `r round_half_up(mean(az_gm_maricopa$pct_nhl_ai_an),0)`% 
(median = `r round_half_up(median(az_gm_maricopa$pct_nhl_ai_an),0)`%) of people 
considered themselves American Indian or Alaskan Native, an average of `r round_half_up(mean(az_gm_maricopa$pct_nhl_asian),0)`% 
(median = `r round_half_up(median(az_gm_maricopa$pct_nhl_asian),0)`%) of people 
considered themselves Asian, fewer than 1% of people considered themselves 
Native Hawaiian or Pacific Islander, and an average of `r round_half_up(mean(az_gm_maricopa$pct_nhl_ao),0)`% 
(median = `r round_half_up(median(az_gm_maricopa$pct_nhl_ao),0)`%) of people 
considered themselves some other race or two or three or more races.

### South Carolina

In those zip codes which lost polling locations (n = `r nrow(sc_demo_lost)`), 
an average of `r round_half_up(mean(sc_demo_lost$pct_hl),0)`% (median = 
`r round_half_up(median(sc_demo_lost$pct_hl),0)`%) of people considered 
themselves Hispanic or Latino, an average of `r round_half_up(mean(sc_demo_lost$pct_nhl_white),0)`% 
(median = `r round_half_up(median(sc_demo_lost$pct_nhl_white),0)`%) of people 
considered themselves White, an average of  `r round_half_up(mean(sc_demo_lost$pct_nhl_black),0)`% 
(median = `r round_half_up(median(sc_demo_lost$pct_nhl_black),0)`%) of people 
considered themselves Black or African-American, fewer than 1% of people 
considered themselves American Indian or Alaskan Native or Native Hawaiian or 
Pacific Islander, an average of `r round_half_up(mean(sc_demo_lost$pct_nhl_asian),0)`% 
(median = `r round_half_up(median(sc_demo_lost$pct_nhl_asian),0)`%) of people 
considered themselves Asian, and an average of `r round_half_up(mean(sc_demo_lost$pct_nhl_ao),0)`% 
(median = `r round_half_up(median(sc_demo_lost$pct_nhl_ao),0)`%) of people 
considered themselves some other race or two or three or more races.

In those zip codes which gained or maintained polling locations (n = `r nrow(sc_demo_gm)`), 
an average of `r round_half_up(mean(sc_demo_gm$pct_hl),0)`% (median = `r round_half_up(median(sc_demo_gm$pct_hl),0)`%) 
of people considered themselves Hispanic or Latino, an average of `r round_half_up(mean(sc_demo_gm$pct_nhl_white),0)`% 
(median = `r round_half_up(median(sc_demo_gm$pct_nhl_white),0)`%) of people 
considered themselves White, an average of `r round_half_up(mean(sc_demo_gm$pct_nhl_black),0)`% 
(median = `r round_half_up(median(sc_demo_gm$pct_nhl_black),0)`%) of people 
considered themselves Black or African-American, fewer than 1% of people 
considered themselves American Indian or Alaskan Native or Native Hawaiian or 
Pacific Islander, an average of `r round_half_up(mean(sc_demo_gm$pct_nhl_asian),0)`% 
(median = `r round_half_up(median(sc_demo_gm$pct_nhl_asian),0)`%) of people 
considered themselves Asian, and an average of `r round_half_up(mean(sc_demo_gm$pct_nhl_ao),0)`% 
(median = `r round_half_up(median(sc_demo_gm$pct_nhl_ao),0)`%) of people 
considered themselves some other race or two or three or more races.

*Richland County*

In those zip codes which lost polling locations within Richland County 
(n = `r nrow(sc_l_rland)`), an average of `r round_half_up(mean(sc_l_rland$pct_hl),0)`% 
(median = `r round_half_up(median(sc_l_rland$pct_hl),0)`%) of people considered 
themselves Hispanic or Latino, an average of `r round_half_up(mean(sc_l_rland$pct_nhl_white),0)`% 
(median = `r round_half_up(median(sc_l_rland$pct_nhl_white),0)`%) of people 
considered themselves White, an average of `r round_half_up(mean(sc_l_rland$pct_nhl_black),0)`% 
(median = `r round_half_up(median(sc_l_rland$pct_nhl_black),0)`%) of people 
considered themselves Black or African-American, fewer than 1% of people 
considered themselves American Indian or Alaskan Native or Native Hawaiian or 
Pacific Islander, an average of `r round_half_up(mean(sc_l_rland$pct_nhl_asian),0)`% 
(median = `r round_half_up(median(sc_l_rland$pct_nhl_asian),0)`%) of people 
considered themselves Asian, and an average of `r round_half_up(mean(sc_l_rland$pct_nhl_ao),0)`% 
(median = `r round_half_up(median(sc_l_rland$pct_nhl_ao),0)`%) of people 
considered themselves some other race or two or three or more races.

\newpage

In those zip codes which gained or maintained polling locations within 
Richland County (n = `r nrow(sc_gm_rland)`), an average of `r round_half_up(mean(sc_gm_rland$pct_hl),0)`% 
(median = `r round_half_up(median(sc_gm_rland$pct_hl),0)`%) of people considered 
themselves Hispanic or Latino, an average of `r round_half_up(mean(sc_gm_rland$pct_nhl_white),0)`% (median = 
`r round_half_up(median(sc_gm_rland$pct_nhl_white),0)`%) of people considered 
themselves White, an average of `r round_half_up(mean(sc_gm_rland$pct_nhl_black),0)`% 
(median = `r round_half_up(median(sc_gm_rland$pct_nhl_black),0)`%) of people 
considered themselves Black or African-American, fewer than 1% of people 
considered themselves American Indian or Alaskan Native or Native Hawaiian or 
Pacific Islander, an average of `r round_half_up(mean(sc_gm_rland$pct_nhl_asian),0)`% 
(median = `r round_half_up(median(sc_gm_rland$pct_nhl_asian),0)`%) of people 
considered themselves Asian, and an average of `r round_half_up(mean(sc_gm_rland$pct_nhl_ao),0)`% 
(median = `r round_half_up(median(sc_gm_rland$pct_nhl_ao),0)`%) of people 
considered themselves some other race or two or three or more races.

## Polling location changes, demographics, and COVID-19 {-} 

```{r covid, echo=FALSE, message=FALSE}

az_demo_covid_df <- read_rds(files$az_demo_covid_clean) %>%
  mutate_at(vars(covid_cat, polling_cat), as.factor) %>%
  select(zipcode, covid_cat, polling_cat, 
         confirmed_case_count, pct_hl, pct_nhl, pct_nhl_white, 
         pct_nhl_black, pct_nhl_ai_an, pct_nhl_asian, pct_nhl_nhi_pi,
         pct_nhl_ao, county) %>%
  mutate(confirmed_case_count = as.numeric(confirmed_case_count))

az_dc_lost <- az_demo_covid_df %>%
  filter(covid_cat == "Lost Polling Places")

az_dc_gm <- az_demo_covid_df %>%
  filter(covid_cat == "Gained or Maintained Polling Places")

az_dc_s <- az_demo_covid_df %>%
  filter(covid_cat == "COVID Data Suppressed Pending Tribal Approval")

az_dc_sgm <- az_demo_covid_df %>%
  filter(covid_cat == "COVID Data Suppressed Pending Tribal Approval" &
           polling_cat == "Gained or Maintained Polling Places")

az_dc_sl  <- az_demo_covid_df %>%
  filter(covid_cat == "COVID Data Suppressed Pending Tribal Approval" &
           polling_cat == "Lost Polling Places")

az_dc_nd <- az_demo_covid_df %>%
  filter(covid_cat == "No COVID Data Reported")

# maricopa county

az_dc_ml <- az_demo_covid_df %>%
  filter(county == "Maricopa County" & covid_cat == "Lost Polling Places")

az_dc_mgm <- az_demo_covid_df %>%
  filter(county == "Maricopa County" & 
           covid_cat == "Gained or Maintained Polling Places")

az_dc_ms <- az_demo_covid_df %>%
  filter(county == "Maricopa County" & 
           covid_cat == "COVID Data Suppressed Pending Tribal Approval")

az_dc_msg <- az_demo_covid_df %>%
  filter(county == "Maricopa County" &
           covid_cat == "COVID Data Suppressed Pending Tribal Approval" &
           polling_cat == "Gained or Maintained Polling Places")

az_dc_msl  <- az_demo_covid_df %>%
  filter(county == "Maricopa County" &
           covid_cat == "COVID Data Suppressed Pending Tribal Approval" &
           polling_cat == "Lost Polling Places")

az_dc_mnd <- az_demo_covid_df %>%
  filter(county == "Maricopa County" & 
           covid_cat == "No COVID Data Reported")

# south carolina

sc_demo_covid_df <- read_rds(files$sc_demo_covid_clean) %>%
  mutate_at(vars(delta_cat), as.factor) %>%
  select(zipcode, cases, delta_cat, pct_hl, pct_nhl, pct_nhl_white, 
         pct_nhl_black, pct_nhl_ai_an, pct_nhl_asian, pct_nhl_nhi_pi,
         pct_nhl_ao, county)

sc_dc_lost <- sc_demo_covid_df %>%
  filter(delta_cat == "Lost Polling Places")

sc_dc_gm <- sc_demo_covid_df %>%
  filter(delta_cat == "Gained or Maintained Polling Places")

# richland county

sc_dc_rl <- sc_demo_covid_df %>%
  filter(county == "Richland" & delta_cat == "Lost Polling Places")

sc_dc_rgm <- sc_demo_covid_df %>%
  filter(county == "Richland" & 
           delta_cat == "Gained or Maintained Polling Places")
# done.
```

### Arizona

In those zip codes with complete COVID-19 data by zip code which lost polling 
locations in the 2020 PPE (n = `r nrow(az_dc_lost)`), an average of `r round_half_up(mean(az_dc_lost$pct_hl),0)`% 
(median = `r round_half_up(median(az_dc_lost$pct_hl),0)`%) of people considered 
themselves Hispanic or Latino, an average of `r round_half_up(mean(az_dc_lost$pct_nhl_white),0)`% 
(median = `r round_half_up(median(az_dc_lost$pct_nhl_white),0)`%) of people 
considered themselves White, an average of `r round_half_up(mean(az_dc_lost$pct_nhl_black),0)`% 
(median = `r round_half_up(median(az_dc_lost$pct_nhl_black),0)`%) of people 
considered themselves Black or African-American, an average of `r round_half_up(mean(az_dc_lost$pct_nhl_ai_an),0)`% 
(median = `r round_half_up(median(az_dc_lost$pct_nhl_ai_an),0)`%) of people 
considered themselves American Indian or Alaskan Native, an average of  `r round_half_up(mean(az_dc_lost$pct_nhl_asian),0)`% 
(median = `r round_half_up(median(az_dc_lost$pct_nhl_asian),0)`%) of people 
considered themselves Asian, fewer than 1% of people considered themselves 
Native Hawaiian or Pacific Islander, and an average of `r round_half_up(mean(az_dc_lost$pct_nhl_ao),0)`% 
(median = `r round_half_up(median(az_dc_lost$pct_nhl_ao),0)`%) of people 
considered themselves some other race or two or three or more races. These zip 
codes had between `r min(az_dc_lost$confirmed_case_count)` - `r max(az_dc_lost$confirmed_case_count)`
cases of COVID-19.

In those zip codes with complete COVID-19 data by zip code which gained or 
maintained polling locations in the 2020 PPE (n = `r nrow(az_dc_gm)`), an average of 
`r round_half_up(mean(az_dc_gm$pct_hl),0)`% (median = `r round_half_up(median(az_dc_gm$pct_hl),0)`%) 
of people considered themselves Hispanic or Latino, an average of `r round_half_up(mean(az_dc_gm$pct_nhl_white),0)`% 
(median = `r round_half_up(median(az_dc_gm$pct_nhl_white),0)`%) of people 
considered themselves White, an average of `r round_half_up(mean(az_dc_gm$pct_nhl_black),0)`% 
(median = `r round_half_up(median(az_dc_gm$pct_nhl_black),0)`%) of people 
considered themselves Black or African-American, an average of  `r round_half_up(mean(az_dc_gm$pct_nhl_ai_an),0)`% 
(median = `r round_half_up(median(az_dc_gm$pct_nhl_ai_an),0)`%) of people 
considered themselves American Indian or Alaskan Native, an average of 
`r round_half_up(mean(az_dc_gm$pct_nhl_asian),0)`% (median = `r round_half_up(median(az_dc_gm$pct_nhl_asian),0)`%) 
of people considered themselves Asian, fewer than 1% of people considered 
themselves Native Hawaiian or Pacific Islander, and an average of `r round_half_up(mean(az_dc_gm$pct_nhl_ao),0)`% 
(median = `r round_half_up(median(az_dc_gm$pct_nhl_ao),0)`%) of people considered 
themselves some other race or two or three or more races. These zip codes had 
between `r min(az_dc_gm$confirmed_case_count)` - `r max(az_dc_gm$confirmed_case_count)`
cases of COVID-19.

In those zip codes with COVID-19 data pending tribal approval which lost polling 
locations in the 2020 PPE (n = `r nrow(az_dc_sl)`), an average of 
`r round_half_up(mean(az_dc_sl$pct_hl),0)`% (median = 
`r round_half_up(median(az_dc_sl$pct_hl),0)`%) of people considered 
themselves Hispanic or Latino, an average of `r round_half_up(mean(az_dc_sl$pct_nhl_white),0)`% 
(median = `r round_half_up(median(az_dc_sl$pct_nhl_white),0)`%) of people 
considered themselves White, fewer than 1% of people considered themselves Black 
or African-American, an average of `r round_half_up(mean(az_dc_sl$pct_nhl_ai_an),0)`% 
(median = `r round_half_up(median(az_dc_sl$pct_nhl_ai_an),0)`%) of people 
considered themselves American Indian or Alaskan Native, fewer than 1% of people 
considered themselves Asian or Native Hawaiian or Pacific Islander, and an 
average of `r round_half_up(mean(az_dc_sl$pct_nhl_ao),0)`% 
(median = `r round_half_up(median(az_dc_sl$pct_nhl_ao),0)`%) of people 
considered themselves some other race or two or three or more races. These zip 
codes had an average of `r mean(az_dc_sl$confirmed_case_count)` case of COVID-19.

In those zip codes with COVID-19 data pending tribal approval which gained or 
maintained polling locations in the 2020 PPE (n = `r nrow(az_dc_sgm)`), an 
average of `r round_half_up(mean(az_dc_sgm$pct_hl),0)`% 
(median = `r round_half_up(median(az_dc_sgm$pct_hl),0)`%) of people considered 
themselves Hispanic or Latino, an average of `r round_half_up(mean(az_dc_sgm$pct_nhl_white),0)`% 
(median = `r round_half_up(median(az_dc_sgm$pct_nhl_white),0)`%) of people 
considered themselves White, fewer than 1% of people considered themselves Black 
or African-American, an average of  `r round_half_up(mean(az_dc_sgm$pct_nhl_ai_an),0)`% 
(median = `r round_half_up(median(az_dc_sgm$pct_nhl_ai_an),0)`%) of people 
considered themselves American Indian or Alaskan Native, fewer than 1% of people 
considered themselves Asian, Native Hawaiian or Pacific Islander, some other 
race or two or three or more races. These zip codes had an average of 
`r mean(az_dc_sgm$confirmed_case_count)` case of COVID-19.

All zip codes without complete COVID-19 data(n = `r nrow(az_dc_nd)`), gained or 
maintained polling locations in the 2020 PPE. An average of  `r round_half_up(mean(az_dc_nd$pct_hl),0)`% 
(median = `r round_half_up(median(az_dc_nd$pct_hl),0)`%) of people considered 
themselves Hispanic or Latino, an average of `r round_half_up(mean(az_dc_nd$pct_nhl_white),0)`% 
(median = `r round_half_up(median(az_dc_nd$pct_nhl_white),0)`%) of people 
considered themselves White, fewer than 1% of people considered themselves Black 
or African-American, an average of `r round_half_up(mean(az_dc_nd$pct_nhl_ai_an),0)`% 
(median = `r round_half_up(median(az_dc_nd$pct_nhl_ai_an),0)`%) of people 
considered themselves American Indian or Alaskan Native, an average of `r round_half_up(mean(az_dc_nd$pct_nhl_asian),0)`% 
(median = `r round_half_up(median(az_dc_nd$pct_nhl_asian),0)`%) of people 
considered themselves Asian, fewer than 1% of people considered themselves 
Native Hawaiian or Pacific Islander, and an average of `r round_half_up(mean(az_dc_nd$pct_nhl_ao),0)`% 
(median = `r round_half_up(median(az_dc_nd$pct_nhl_ao),0)`%) of people 
considered themselves some other race or two or three or more races.

\newpage

*Maricopa County*

In those zip codes within Maricopa County with complete COVID-19 data by zip code 
which lost polling locations in the 2020 PPE (n = `r nrow(az_dc_ml)`), an average 
of `r round_half_up(mean(az_dc_ml$pct_hl),0)`% (median = `r round_half_up(median(az_dc_ml$pct_hl),0)`%) 
of people considered themselves Hispanic or Latino, an average of 
`r round_half_up(mean(az_dc_ml$pct_nhl_white),0)`% (median = `r round_half_up(median(az_dc_ml$pct_nhl_white),0)`%) 
of people considered themselves White, an average of `r round_half_up(mean(az_dc_ml$pct_nhl_black),0)`% 
(median = `r round_half_up(median(az_dc_ml$pct_nhl_black),0)`%) of people 
considered themselves Black or African-American, an average of `r round_half_up(mean(az_dc_ml$pct_nhl_ai_an),0)`% 
(median = `r round_half_up(median(az_dc_ml$pct_nhl_ai_an),0)`%) of people 
considered themselves American Indian or Alaskan Native, an average of `r round_half_up(mean(az_dc_ml$pct_nhl_asian),0)`% 
(median = `r round_half_up(median(az_dc_ml$pct_nhl_asian),0)`%) of people 
considered themselves Asian, fewer than 1% of people considered themselves 
Native Hawaiian or Pacific Islander, and an average of `r round_half_up(mean(az_dc_ml$pct_nhl_ao),0)`% 
(median = `r round_half_up(median(az_dc_ml$pct_nhl_ao),0)`%) of people considered
themselves some other race or two or three or more races. These zip codes had 
between `r min(az_dc_ml$confirmed_case_count)` - `r max(az_dc_ml$confirmed_case_count)`
cases of COVID-19.

In those zip codes within Maricopa County with complete COVID-19 data by zip code 
which gained or maintained polling locations in the 2020 PPE (n = `r nrow(az_dc_mgm)`), 
an average of `r round_half_up(mean(az_dc_mgm$pct_hl),0)`% (median = `r round_half_up(median(az_dc_mgm$pct_hl),0)`%) 
of people considered themselves Hispanic or Latino, an average of `r round_half_up(mean(az_dc_mgm$pct_nhl_white),0)`% 
(median = `r round_half_up(median(az_dc_mgm$pct_nhl_white),0)`%) of people 
considered themselves White, an average of `r round_half_up(mean(az_dc_mgm$pct_nhl_black),0)`% 
(median = `r round_half_up(median(az_dc_mgm$pct_nhl_black),0)`%) of people 
considered themselves Black or African-American, an average of  `r round_half_up(mean(az_dc_mgm$pct_nhl_ai_an),0)`% 
(median = `r round_half_up(median(az_dc_mgm$pct_nhl_ai_an),0)`%) of people 
considered themselves American Indian or Alaskan Native, an average of `r round_half_up(mean(az_dc_mgm$pct_nhl_asian),0)`% 
(median = `r round_half_up(median(az_dc_mgm$pct_nhl_asian),0)`%) of people 
considered themselves Asian, fewer than 1% of people considered themselves 
Native Hawaiian or Pacific Islander, and an average of  `r round_half_up(mean(az_dc_mgm$pct_nhl_ao),0)`% 
(median = `r round_half_up(median(az_dc_mgm$pct_nhl_ao),0)`%) of people 
considered themselves some other race or two or three or more races. These zip 
codes had between
`r min(az_dc_mgm$confirmed_case_count)` - `r max(az_dc_mgm$confirmed_case_count)`
cases of COVID-19.

There were no zip codes within within Maricopa County which had any COVID-19 
data pending tribal approval nor were there any zip codes for which COVID-19
data were unavailable.

### South Carolina

In those zip codes with complete COVID-19 data by zip code which lost polling 
locations in the 2020 PPE (n = `r nrow(sc_dc_lost)`), an average of `r round_half_up(mean(sc_dc_lost$pct_hl),0)`% 
(median = `r round_half_up(median(sc_dc_lost$pct_hl),0)`%) of people considered 
themselves Hispanic or Latino, an average of `r round_half_up(mean(sc_dc_lost$pct_nhl_white),0)`% 
(median = `r round_half_up(median(sc_dc_lost$pct_nhl_white),0)`%) of people 
considered themselves White, an average of `r round_half_up(mean(sc_dc_lost$pct_nhl_black),0)`% 
(median = `r round_half_up(median(sc_dc_lost$pct_nhl_black),0)`%) of people 
considered themselves Black or African-American, fewer than 1% of people 
considered themselves American Indian or Alaskan Native or Native Hawaiian or 
Pacific Islander, an average of  `r round_half_up(mean(sc_dc_lost$pct_nhl_asian),0)`% 
(median = `r round_half_up(median(sc_dc_lost$pct_nhl_asian),0)`%) of people 
considered themselves Asian, and an average of `r round_half_up(mean(sc_dc_lost$pct_nhl_ao),0)`% 
(median = `r round_half_up(median(sc_dc_lost$pct_nhl_ao),0)`%) of people 
considered themselves some other race or two or three or more races. These zip 
codes had between 
`r comma(min(sc_dc_lost$cases))` - `r comma(max(sc_dc_lost$cases))` cases of 
COVID-19.

In those zip codes with complete COVID-19 data by zip code which gained or 
maintained polling locations in the 2020 PPE (n = `r nrow(sc_dc_gm)`), an 
average of `r round_half_up(mean(sc_dc_gm$pct_hl),0)`% 
(median = `r round_half_up(median(sc_dc_gm$pct_hl),0)`%) of people considered 
themselves Hispanic or Latino, an average of `r round_half_up(mean(sc_dc_gm$pct_nhl_white),0)`% 
(median = `r round_half_up(median(sc_dc_gm$pct_nhl_white),0)`%) of people 
considered themselves White, an average of `r round_half_up(mean(sc_dc_gm$pct_nhl_black),0)`% 
(median = `r round_half_up(median(sc_dc_gm$pct_nhl_black),0)`%) of people 
considered themselves Black or African-American, fewer than 1% of people 
considered themselves American Indian or Alaskan Native or Native Hawaiian or 
Pacific Islander, an average of `r round_half_up(mean(sc_dc_gm$pct_nhl_asian),0)`% 
(median = `r round_half_up(median(sc_dc_gm$pct_nhl_asian),0)`%) of people 
considered themselves Asian, and an average of `r round_half_up(mean(sc_dc_gm$pct_nhl_ao),0)`% 
(median = `r round_half_up(median(sc_dc_gm$pct_nhl_ao),0)`%) of people 
considered themselves some other race or two or three or more races. These zip 
codes had between `r comma(min(sc_dc_gm$cases))` - `r comma(max(sc_dc_gm$cases))` 
cases of COVID-19.

*Richland County*

In those zip codes within Richland County with complete COVID-19 data by zip code 
which lost polling locations in the 2020 PPE (n = `r nrow(sc_dc_rl)`), an average 
of `r round_half_up(mean(sc_dc_rl$pct_hl),0)`% (median = `r round_half_up(median(sc_dc_rl$pct_hl),0)`%) 
of people considered themselves Hispanic or Latino, an average of 
`r round_half_up(mean(sc_dc_rl$pct_nhl_white),0)`% (median = `r round_half_up(median(sc_dc_rl$pct_nhl_white),0)`%) 
of people considered themselves White, an average of `r round_half_up(mean(sc_dc_rl$pct_nhl_black),0)`% 
(median = `r round_half_up(median(sc_dc_rl$pct_nhl_black),0)`%) of people 
considered themselves Black or African-American, fewer than 1% of people 
considered themselves American Indian or Alaskan Native or Native Hawaiian or 
Pacific Islander, an average of `r round_half_up(mean(sc_dc_rl$pct_nhl_asian),0)`% 
(median = `r round_half_up(median(sc_dc_rl$pct_nhl_asian),0)`%) of people 
considered themselves Asian, and an average of `r round_half_up(mean(sc_dc_rl$pct_nhl_ao),0)`% 
(median = `r round_half_up(median(sc_dc_rl$pct_nhl_ao),0)`%) of people considered
themselves some other race or two or three or more races. These zip codes had an
average of `r comma(mean(sc_dc_rl$cases))` cases of COVID-19.

\newpage

In those zip codes within Richland County with complete COVID-19 data by zip code 
which gained or maintained polling locations in the 2020 PPE (n = `r nrow(sc_dc_rgm)`), 
an average of `r round_half_up(mean(sc_dc_rgm$pct_hl),0)`% 
(median = `r round_half_up(median(sc_dc_rgm$pct_hl),0)`%) of people considered 
themselves Hispanic or Latino, an average of `r round_half_up(mean(sc_dc_rgm$pct_nhl_white),0)`% 
(median = `r round_half_up(median(sc_dc_rgm$pct_nhl_white),0)`%) of people 
considered themselves White, an average of `r round_half_up(mean(sc_dc_rgm$pct_nhl_black),0)`% 
(median = `r round_half_up(median(sc_dc_rgm$pct_nhl_black),0)`%) of people 
considered themselves Black or African-American, fewer than 1% of people 
considered themselves American Indian or Alaskan Native or Native Hawaiian or 
Pacific Islander, an average of `r round_half_up(mean(sc_dc_rgm$pct_nhl_asian),0)`% 
(median = `r round_half_up(median(sc_dc_rgm$pct_nhl_asian),0)`%) of people 
considered themselves Asian, and an average of `r round_half_up(mean(sc_dc_rgm$pct_nhl_ao),0)`% 
(median = `r round_half_up(median(sc_dc_rgm$pct_nhl_ao),0)`%) of people 
considered themselves some other race or two or three or more races. These zip 
codes had an average of `r comma(mean(sc_dc_rgm$cases))` cases of COVID-19.

## Methodology

This report examines polling location data from `r nrow(ncountiesaz)`
or `r round_half_up(sum((nrow(ncountiesaz)/15)*100),0)`% of the counties in 
Arizona and `r nrow(ncountiessc)` or `r round_half_up(sum((nrow(ncountiessc)/46)*100),0)`% 
of the counties in South Carolina. Pima County, AZ has not been included since 
the data come from the Voter Information Project and Pima County does not 
participate in its activities. Data were reviewed for uniqueness, completeness, 
and accuracy of information on zip codes, county, and number of polling locations. 
Polling locations are known to change just before an election so this data 
should not be viewed as a definitive list of the actual polling locations used 
on election day.

Demographic data were obtained from the 2014-2018 American Community Survey's 
B03002 table and groups were categorized as either Hispanic or Latino, and
not Hispanic or Latino by specific group: White alone, Black or African American 
Alone, American Indian and Alaska Native alone, Asian alone,Native Hawaiian and 
Other Pacific Islander alone, and All other races alone (Including Some other 
race alone and Two or more races alone).

Zip code and county data were obtained from unitedstateszipcodes.org. 
While many zip codes map directly to specific counties, the relationship between 
these two organizational systems is not always 1:1 and is subject to change.

Data on COVID-19 cases in Arizona was obtained from the Arizona Department 
of Health Services while data on cases in South Carolina was obtained
though the New York Times covid-19 repository on GitHub. Data from these sources
was combined with demographic and zip code data. Data on COVID-19 cases was only 
kept for those zip codes for which we also had polling location data and in 
Arizona, was categorized to denote where data were missing or suppressed pending 
tribal approval as noted in the original data. Data are current as of as 
of `r format(Sys.time(), '%d %B, %Y')`.

## Acknowledgements {-} 

We are grateful to Sarah Berlin at Democracy Works for collecting data from the 
Voter Information Project.