---
title: 'Polling Locations in Arizona and South Carolina: 2016 vs 2020'
author: "Jessica Randall"
date: "Last compiled `r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    toc: false
    number_sections: false
---

```{r setup, message = FALSE, include= FALSE, echo=FALSE}

pacman::p_load(tidyverse, knitr, janitor, tinytex, scales)
# remove scientific notation from output
options(scipen = 999)

files <- list(
  az_2016_clean = here::here("write/input/az_2016_clean.rds"),
  az_2016_freq_clean = here::here("write/input/az_2016_freq_clean.rds"),
  az_2020_clean = here::here("write/input/az_2020_clean.rds"),
  az_2020_freq_clean = here::here("write/input/az_2020_freq_clean.rds"),
  az_demo_clean = here::here("write/input/az_demo_clean.rds"),
  az_demo_covid_clean = here::here("write/input/az_demo_covid_clean.rds"),
  sc_2016_clean = here::here("write/input/sc_2016_clean.rds"),
  sc_2016_freq_clean = here::here("write/input/sc_2016_freq_clean.rds"),
  sc_2020_clean = here::here("write/input/sc_2020_clean.rds"),
  sc_2020_freq_clean = here::here("write/input/sc_2020_freq_clean.rds"),
  sc_demo_clean = here::here("write/input/sc_demo_clean.rds"), 
  sc_demo_covid_clean = here::here("write/input/sc_demo_covid_clean.rds")
)

```

## Polling locations open in 2016 and 2020 {-} 

```{r nlocationsa, message=FALSE, echo=FALSE}

az_2016_df <- read_rds(files$az_2016_clean)
az_2020_df <- read_rds(files$az_2020_clean)

sc_2016_df <- read_rds(files$sc_2016_clean)
sc_2020_df <- read_rds(files$sc_2020_clean)

```

### Arizona

Overall, Arizona had `r nrow(az_2016_df)` open polling locations statewide at the 2016
primary Presidential election (PPE). During the 2020 PPE there were `r nrow(az_2020_df)` 
open polling locations. This is a difference of `r sum((nrow(az_2020_df) - nrow(az_2016_df)))` 
more or a `r round_half_up((sum((nrow(az_2020_df) - nrow(az_2016_df)))/as.numeric(nrow(az_2016_df))*100),1)`% 
increase in polling locations.

### South Carolina

Overall, South Carolina had `r nrow(sc_2016_df)` open polling locations statewide at the 2016
PPE. During the 2020 PPE there were `r nrow(sc_2020_df)` open polling locations. 
This is a difference of `r abs(sum((nrow(sc_2020_df) - nrow(sc_2016_df))))` fewer 
or a `r abs(round_half_up((sum((nrow(sc_2020_df) - nrow(sc_2016_df)))/as.numeric(nrow(sc_2016_df))*100),1))`% 
decrease in polling locations.

## Changes polling locations between 2016 and 2020 {-} 

```{r nlocationsb, message=FALSE, echo=FALSE}
# AZ
az_zips_freq_2016 <- read_rds(files$az_2016_freq_clean)
az_zips_freq_2020 <- read_rds(files$az_2020_freq_clean)

n_places_az <- full_join(az_zips_freq_2020, az_zips_freq_2016, by = "zipcode") %>%
  replace_na(list(n_pp_2016 = 0, n_pp_2020 = 0)) %>%
  mutate(delta_n_places = as.numeric(n_pp_2020 - n_pp_2016),
         delta_cat = case_when(
           delta_n_places == 0 ~ "Maintained Polling Locations",
           delta_n_places < 0 ~ "Lost Polling Places",
           delta_n_places > 0 ~ "Gained Polling Locations")) %>%
  arrange(delta_n_places) %>%
  select(zipcode, n_pp_2020, n_pp_2016, delta_n_places, delta_cat)

# how many lost, and how many gained, maintained?
az_table <- n_places_az %>%
  group_by(delta_cat) %>%
  summarise(.groups = "keep", 
            count = n(),
            percent_of_zips = ((count/nrow(n_places_az))*100))

# SC
sc_zips_freq_2016 <- read_rds(files$sc_2016_freq_clean)
sc_zips_freq_2020 <- read_rds(files$sc_2020_freq_clean)

# which zip codes saw an increase, decrease, or maintenance in polling places?
n_places_sc <- full_join(sc_zips_freq_2020, sc_zips_freq_2016, by = "zipcode") %>%
  replace_na(list(n_pp_2016 = 0, n_pp_2020 = 0)) %>%
  mutate(delta_n_places = as.numeric(n_pp_2020 - n_pp_2016),
         delta_cat = case_when(
           delta_n_places == 0 ~ "Maintained Polling Locations",
           delta_n_places < 0 ~ "Lost Polling Places",
           delta_n_places > 0 ~ "Gained Polling Locations")) %>%
  filter(zipcode != "SC") %>%
  arrange(delta_n_places) %>%
  select(zipcode, n_pp_2020, n_pp_2016, delta_n_places, delta_cat)

# how many lost, and how many gained, maintained?
sc_table <- n_places_sc %>%
  group_by(delta_cat) %>%
  summarise(.groups = "keep", 
            count = n(),
            percent_of_zips = (count/nrow(n_places_sc)*100))

```

### Arizona 

Between the 2016 and 2020 PPEs, `r az_table$count[1]` zip codes, or 
`r round_half_up(az_table$percent_of_zips[1],1)`% of zip codes in Arizona with 
open polling locations during the 2016 PPE gained polling locations during the 
2020 PPE while `r az_table$count[2]` zip codes, 
or `r round_half_up(az_table$percent_of_zips[2],1)`%, maintained the same
number of polling locations, and `r az_table$count[3]` zip codes, 
or `r round_half_up(az_table$percent_of_zips[3],1)`% lost polling locations
during the 2020 PPE.

### South Carolina 
  
Between the 2016 and 2020 PPEs, `r sc_table$count[1]` zip codes, or 
`r round_half_up(sc_table$percent_of_zips[1],1)`% of zip codes in Arizona with 
open polling locations during the 2016 PPE gained polling locations during the 
2020 PPE while `r sc_table$count[2]` zip codes, 
or `r round_half_up(sc_table$percent_of_zips[2],1)`%, maintained the same
number of polling locations, and `r sc_table$count[3]` zip codes, 
or `r round_half_up(sc_table$percent_of_zips[3],1)`% lost polling locations
during the 2020 PPE.
 
## Polling location changes and demographics {-} 

```{r demographics, message=FALSE, echo=FALSE}

az <- read_rds(files$az_demo_clean)

az_lost <- az %>%
  filter(delta_cat == "Lost Polling Locations")
az_g <- az %>%
  filter(delta_cat == "Gained Polling Locations")
az_m <- az %>%
  filter(delta_cat == "Maintained Polling Locations")

az_lost_maricopa <- az_lost %>%
  filter(county == "Maricopa County")
az_g_maricopa <- az_g %>%
  filter(county == "Maricopa County")
az_m_maricopa <- az_m %>%
  filter(county == "Maricopa County")

ncountiesaz <- az %>%
  group_by(county) %>%
  distinct(county) %>%
  arrange(county)

sc <- read_rds(files$sc_demo_clean)

sc_lost <- sc %>%
  filter(delta_cat == "Lost Polling Locations")
sc_g <- sc %>%
  filter(delta_cat == "Gained Polling Locations")
sc_m <- sc %>%
  filter(delta_cat == "Maintained Polling Locations")

sc_lost_rland <- sc_lost %>%
  filter(county == "Richland")
sc_g_rland <- sc_g %>%
  filter(county == "Richland")
sc_m_rland <- sc_m %>%
  filter(county == "Richland")

ncountiessc <- sc %>%
  group_by(county) %>%
  distinct(county) %>%
  arrange(county)

```

### Arizona

In those zip codes which gained polling locations (n = `r nrow(az_g)`),
`r comma(sum(az_g$total_hl))`people considered themselves Hispanic or Latino, 
`r comma(sum(az_g$nhl_white))` people considered themselves White, 
`r comma(sum(az_g$nhl_black))` people considered themselves Black or African-American, 
`r comma(sum(az_g$nhl_ai_an))` people considered themselves American Indian or Alaskan 
Native, `r comma(sum(az_g$nhl_asian))` people considered themselves Asian, 
`r comma(sum(az_g$nhl_nhi_pi))` people considered themselves Native Hawaiian or Pacific 
Islander, and `r comma(sum(az_g$pct_nhl_ao))`people considered themselves some other 
race or two or three or more races.

In those zip codes which maintained polling locations (n = `r nrow(az_m)`),
`r comma(sum(az_m$total_hl))`people considered themselves Hispanic or Latino, 
`r comma(sum(az_m$nhl_white))` people considered themselves White, 
`r comma(sum(az_m$nhl_black))` people considered themselves Black or African-American, 
`r comma(sum(az_m$nhl_ai_an))` people considered themselves American Indian or Alaskan 
Native, `r comma(sum(az_m$nhl_asian))` people considered themselves Asian, 
`r comma(sum(az_m$nhl_nhi_pi))` people considered themselves Native Hawaiian or Pacific 
Islander, and `r comma(sum(az_m$pct_nhl_ao))`people considered themselves some other 
race or two or three or more races.

In those zip codes which maintained polling locations (n = `r nrow(az_lost)`),
`r comma(sum(az_lost$total_hl))`people considered themselves Hispanic or Latino, 
`r comma(sum(az_lost$nhl_white))` people considered themselves White, 
`r comma(sum(az_lost$nhl_black))` people considered themselves Black or African-American, 
`r comma(sum(az_lost$nhl_ai_an))` people considered themselves American Indian or Alaskan 
Native, `r comma(sum(az_lost$nhl_asian))` people considered themselves Asian, 
`r comma(sum(az_lost$nhl_nhi_pi))` people considered themselves Native Hawaiian or Pacific 
Islander, and `r comma(sum(az_lost$pct_nhl_ao))`people considered themselves some other 
race or two or three or more races.

*Maricopa County*

In those zip codes which gained polling locations (n = `r nrow(az_g_maricopa)`),
`r comma(sum(az_g_maricopa$total_hl))`people considered themselves Hispanic or Latino, 
`r comma(sum(az_g_maricopa$nhl_white))` people considered themselves White, 
`r comma(sum(az_g_maricopa$nhl_black))` people considered themselves Black or African-American, 
`r comma(sum(az_g_maricopa$nhl_ai_an))` people considered themselves American Indian or Alaskan 
Native, `r comma(sum(az_g_maricopa$nhl_asian))` people considered themselves Asian, 
`r comma(sum(az_g_maricopa$nhl_nhi_pi))` people considered themselves Native Hawaiian or Pacific 
Islander, and `r comma(sum(az_g_maricopa$pct_nhl_ao))`people considered themselves some other 
race or two or three or more races.

In those zip codes which maintained polling locations (n = `r nrow(az_m_maricopa)`),
`r comma(sum(az_m_maricopa$total_hl))`people considered themselves Hispanic or Latino, 
`r comma(sum(az_m_maricopa$nhl_white))` people considered themselves White, 
`r comma(sum(az_m_maricopa$nhl_black))` people considered themselves Black or African-American, 
`r comma(sum(az_m_maricopa$nhl_ai_an))` people considered themselves American Indian or Alaskan 
Native, `r comma(sum(az_m_maricopa$nhl_asian))` people considered themselves Asian, 
`r comma(sum(az_m_maricopa$nhl_nhi_pi))` people considered themselves Native Hawaiian or Pacific 
Islander, and `r comma(sum(az_m_maricopa$pct_nhl_ao))`people considered themselves some other 
race or two or three or more races.

In those zip codes which maintained polling locations (n = `r nrow(az_lost_maricopa)`),
`r comma(sum(az_lost_maricopa$total_hl))`people considered themselves Hispanic or Latino, 
`r comma(sum(az_lost_maricopa$nhl_white))` people considered themselves White, 
`r comma(sum(az_lost_maricopa$nhl_black))` people considered themselves Black or African-American, 
`r comma(sum(az_lost_maricopa$nhl_ai_an))` people considered themselves American Indian or Alaskan 
Native, `r comma(sum(az_lost_maricopa$nhl_asian))` people considered themselves Asian, 
`r comma(sum(az_lost_maricopa$nhl_nhi_pi))` people considered themselves Native Hawaiian or Pacific 
Islander, and `r comma(sum(az_lost_maricopa$pct_nhl_ao))`people considered themselves some other 
race or two or three or more races.

### South Carolina

In those zip codes which gained polling locations (n = `r nrow(sc_g)`),
`r comma(sum(sc_g$total_hl))`people considered themselves Hispanic or Latino, 
`r comma(sum(sc_g$nhl_white))` people considered themselves White, 
`r comma(sum(sc_g$nhl_black))` people considered themselves Black or African-American, 
`r comma(sum(sc_g$nhl_ai_an))` people considered themselves American Indian or Alaskan 
Native, `r comma(sum(sc_g$nhl_asian))` people considered themselves Asian, 
`r comma(sum(sc_g$nhl_nhi_pi))` people considered themselves Native Hawaiian or Pacific 
Islander, and `r comma(sum(sc_g$pct_nhl_ao))`people considered themselves some other 
race or two or three or more races.

In those zip codes which maintained polling locations (n = `r nrow(sc_m)`),
`r comma(sum(sc_m$total_hl))`people considered themselves Hispanic or Latino, 
`r comma(sum(sc_m$nhl_white))` people considered themselves White, 
`r comma(sum(sc_m$nhl_black))` people considered themselves Black or African-American, 
`r comma(sum(sc_m$nhl_ai_an))` people considered themselves American Indian or Alaskan 
Native, `r comma(sum(sc_m$nhl_asian))` people considered themselves Asian, 
`r comma(sum(sc_m$nhl_nhi_pi))` people considered themselves Native Hawaiian or Pacific 
Islander, and `r comma(sum(sc_m$pct_nhl_ao))`people considered themselves some other 
race or two or three or more races.

In those zip codes which maintained polling locations (n = `r nrow(sc_lost)`),
`r comma(sum(sc_lost$total_hl))`people considered themselves Hispanic or Latino, 
`r comma(sum(sc_lost$nhl_white))` people considered themselves White, 
`r comma(sum(sc_lost$nhl_black))` people considered themselves Black or African-American, 
`r comma(sum(sc_lost$nhl_ai_an))` people considered themselves American Indian or Alaskan 
Native, `r comma(sum(sc_lost$nhl_asian))` people considered themselves Asian, 
`r comma(sum(sc_lost$nhl_nhi_pi))` people considered themselves Native Hawaiian or Pacific 
Islander, and `r comma(sum(sc_lost$pct_nhl_ao))`people considered themselves some other 
race or two or three or more races.

*Richland County*

In those zip codes which gained polling locations (n = `r nrow(sc_g_rland)`),
`r comma(sum(sc_g_rland$total_hl))`people considered themselves Hispanic or Latino, 
`r comma(sum(sc_g_rland$nhl_white))` people considered themselves White, 
`r comma(sum(sc_g_rland$nhl_black))` people considered themselves Black or African-American, 
`r comma(sum(sc_g_rland$nhl_ai_an))` people considered themselves American Indian or Alaskan 
Native, `r comma(sum(sc_g_rland$nhl_asian))` people considered themselves Asian, 
`r comma(sum(sc_g_rland$nhl_nhi_pi))` people considered themselves Native Hawaiian or Pacific 
Islander, and `r comma(sum(sc_g_rland$pct_nhl_ao))`people considered themselves some other 
race or two or three or more races.

In those zip codes which maintained polling locations (n = `r nrow(sc_m_rland)`),
`r comma(sum(sc_m_rland$total_hl))`people considered themselves Hispanic or Latino, 
`r comma(sum(sc_m_rland$nhl_white))` people considered themselves White, 
`r comma(sum(sc_m_rland$nhl_black))` people considered themselves Black or African-American, 
`r comma(sum(sc_m_rland$nhl_ai_an))` people considered themselves American Indian or Alaskan 
Native, `r comma(sum(sc_m_rland$nhl_asian))` people considered themselves Asian, 
`r comma(sum(sc_m_rland$nhl_nhi_pi))` people considered themselves Native Hawaiian or Pacific 
Islander, and `r comma(sum(sc_m_rland$pct_nhl_ao))`people considered themselves some other 
race or two or three or more races.

In those zip codes which maintained polling locations (n = `r nrow(sc_lost_rland)`),
`r comma(sum(sc_lost_rland$total_hl))`people considered themselves Hispanic or Latino, 
`r comma(sum(sc_lost_rland$nhl_white))` people considered themselves White, 
`r comma(sum(sc_lost_rland$nhl_black))` people considered themselves Black or African-American, 
`r comma(sum(sc_lost_rland$nhl_ai_an))` people considered themselves American Indian or Alaskan 
Native, `r comma(sum(sc_lost_rland$nhl_asian))` people considered themselves Asian, 
`r comma(sum(sc_lost_rland$nhl_nhi_pi))` people considered themselves Native Hawaiian or Pacific 
Islander, and `r comma(sum(sc_lost_rland$pct_nhl_ao))`people considered themselves some other 
race or two or three or more races.

## Polling location changes, demographics, and COVID-19 {-} 

```{r covid, echo=FALSE, message=FALSE}

az_covid_df <- read_rds(files$az_demo_covid_clean) %>%
  mutate_at(vars(covid_cat, polling_cat), as.factor) %>%
  select(zipcode, covid_cat, polling_cat, 
         confirmed_case_count, pct_hl, pct_nhl, pct_nhl_white, 
         pct_nhl_black, pct_nhl_ai_an, pct_nhl_asian, pct_nhl_nhi_pi,
         pct_nhl_ao, county) %>%
  mutate(confirmed_case_count = as.numeric(confirmed_case_count))

az_dc_lost <- az_covid_df %>%
  filter(covid_cat == "Lost Polling Locations")
az_dc_g <- az_covid_df %>%
  filter(covid_cat == "Gained Polling Locations")
az_dc_m <- az_covid_df %>%
  filter(covid_cat == "Maintained Polling Locations")

az_dc_s <- az_covid_df %>%
  filter(covid_cat == "COVID Data Suppressed Pending Tribal Approval")
az_dcs_lost <- az_dc_s %>%
  filter(polling_cat == "Lost Polling Locations")
az_dcs_g <- az_dc_s %>%
  filter(polling_cat == "Gained Polling Locations")
az_dcs_m <- az_dc_s %>%
  filter(polling_cat == "Maintained Polling Locations")

az_dc_nd <- az_covid_df %>%
  filter(covid_cat == "No COVID Data Reported")
az_dc_ndg <- az_dc_nd %>%
  filter(polling_cat == "Gained Polling Locations")
az_dc_ndm <- az_dc_nd %>%
  filter(polling_cat == "Maintained Polling Locations")

# maricopa county
az_dclost_maricopa <- az_dc_lost %>%
  filter(county == "Maricopa County")
az_dcg_maricopa <- az_dc_g %>%
  filter(county == "Maricopa County")
az_dcm_maricopa <- az_dc_m %>%
  filter(county == "Maricopa County")
az_dc_ndmaricopa <- az_dc_nd %>%
  filter(covid_cat == "No COVID Data Reported")
az_dcg_ndmaricopa <- az_dc_nd %>%
  filter(polling_cat == "Gained Polling Locations")
az_dcm_ndmaricopa <- az_dc_nd %>%
  filter(polling_cat == "Maintained Polling Locations")

# south carolina

sc_covid_df <- read_rds(files$sc_demo_covid_clean) %>%
  mutate_at(vars(delta_cat), as.factor) %>%
  select(zipcode, cases, delta_cat, pct_hl, pct_nhl, pct_nhl_white, 
         pct_nhl_black, pct_nhl_ai_an, pct_nhl_asian, pct_nhl_nhi_pi,
         pct_nhl_ao, county)

sc_dc_lost <- sc_covid_df %>%
  filter(delta_cat == "Lost Polling Locations")
sc_dc_g <- sc_covid_df %>%
  filter(delta_cat == "Gained Polling Locations")
sc_dc_m <- sc_covid_df %>%
  filter(delta_cat == "Maintained Polling Locations")

# Richland county
sc_dclost_richland <- sc_dc_lost %>%
  filter(county == "Richland")
sc_dcg_richland <- sc_dc_g %>%
  filter(county == "Richland")
sc_dcm_richland <- sc_dc_m %>%
  filter(county == "Richland")

# done.
```

#FIXME: add COVID ranges

### Arizona

In those zip codes with complete COVID-19 data which gained polling locations 
(n = `r nrow(az_dc_g)`),`r comma(sum(az_dc_g$total_hl))`people considered themselves Hispanic 
or Latino, `r comma(sum(az_dc_g$nhl_white))` people considered themselves White, 
`r comma(sum(az_dc_g$nhl_black))` people considered themselves Black or African-American, 
`r comma(sum(az_dc_g$nhl_ai_an))` people considered themselves American Indian or Alaskan 
Native, `r comma(sum(az_dc_g$nhl_asian))` people considered themselves Asian, 
`r comma(sum(az_dc_g$nhl_nhi_pi))` people considered themselves Native Hawaiian or Pacific 
Islander, and `r comma(sum(az_dc_g$pct_nhl_ao))`people considered themselves some other race 
or two or three or more races.

In those zip codes with complete COVID-19 data which maintained polling locations 
(n = `r nrow(az_dc_m)`),`r comma(sum(az_dc_m$total_hl))`people considered themselves Hispanic 
or Latino, `r comma(sum(az_dc_m$nhl_white))` people considered themselves White, 
`r comma(sum(az_dc_m$nhl_black))` people considered themselves Black or African-American, 
`r comma(sum(az_dc_m$nhl_ai_an))` people considered themselves American Indian or Alaskan 
Native, `r comma(sum(az_dc_m$nhl_asian))` people considered themselves Asian, 
`r comma(sum(az_dc_m$nhl_nhi_pi))` people considered themselves Native Hawaiian or Pacific 
Islander, and `r comma(sum(az_dc_m$pct_nhl_ao))`people considered themselves some other race 
or two or three or more races.

In those zip codes with complete COVID-19 data which lost polling locations 
(n = `r nrow(az_dc_lost)`),`r comma(sum(az_dc_lost$total_hl))`people considered themselves Hispanic 
or Latino, `r comma(sum(az_dc_lost$nhl_white))` people considered themselves White, 
`r comma(sum(az_dc_lost$nhl_black))` people considered themselves Black or African-American, 
`r comma(sum(az_dc_lost$nhl_ai_an))` people considered themselves American Indian or Alaskan 
Native, `r comma(sum(az_dc_lost$nhl_asian))` people considered themselves Asian, 
`r comma(sum(az_dc_lost$nhl_nhi_pi))` people considered themselves Native Hawaiian or Pacific 
Islander, and `r comma(sum(az_dc_lost$pct_nhl_ao))`people considered themselves some other race 
or two or three or more races.

In those zip codes in which COVID-19 data are pending tribal approval and 
which gained polling locations (n = `r nrow(az_dcs_g)`),
`r comma(sum(az_dcs_g$total_hl))`people considered themselves Hispanic or Latino, 
`r comma(sum(az_dcs_g$nhl_white))` people considered themselves White, 
`r comma(sum(az_dcs_g$nhl_black))` people considered themselves Black or African-American, 
`r comma(sum(az_dcs_g$nhl_ai_an))` people considered themselves American Indian or 
Alaskan Native, `r comma(sum(az_dcs_g$nhl_asian))` people considered themselves Asian, 
`r comma(sum(az_dcs_g$nhl_nhi_pi))` people considered themselves Native Hawaiian or 
Pacific Islander, and `r comma(sum(az_dcs_g$pct_nhl_ao))`people considered themselves 
some other race or two or three or more races.

In those zip codes in which COVID-19 data are pending tribal approval and 
which gained polling locations (n = `r nrow(az_dcs_m)`),
`r comma(sum(az_dcs_m$total_hl))`people considered themselves Hispanic or Latino, 
`r comma(sum(az_dcs_m$nhl_white))` people considered themselves White, 
`r comma(sum(az_dcs_m$nhl_black))` people considered themselves Black or African-American, 
`r comma(sum(az_dcs_m$nhl_ai_an))` people considered themselves American Indian or 
Alaskan Native, `r comma(sum(az_dcs_m$nhl_asian))` people considered themselves Asian, 
`r comma(sum(az_dcs_m$nhl_nhi_pi))` people considered themselves Native Hawaiian or 
Pacific Islander, and `r comma(sum(az_dcs_m$pct_nhl_ao))`people considered themselves 
some other race or two or three or more races.

In those zip codes in which COVID-19 data are pending tribal approval and 
which gained polling locations (n = `r nrow(az_dcs_lost)`),
`r comma(sum(az_dcs_lost$total_hl))` people considered themselves Hispanic or Latino, 
`r comma(sum(az_dcs_lost$nhl_white))` people considered themselves White, 
`r comma(sum(az_dcs_lost$nhl_black))` people considered themselves Black or African-American, 
`r comma(sum(az_dcs_lost$nhl_ai_an))` people considered themselves American Indian or 
Alaskan Native, `r comma(sum(az_dcs_lost$nhl_asian))` people considered themselves Asian, 
`r comma(sum(az_dcs_lost$nhl_nhi_pi))` people considered themselves Native Hawaiian or 
Pacific Islander, and `r comma(sum(az_dcs_lost$pct_nhl_ao))`people considered themselves 
some other race or two or three or more races.

In those zip codes without COVID-19 data which gained polling locations 
(n = `r nrow(az_dc_ndg)`),`r comma(sum(az_dc_ndg$total_hl))`people considered themselves Hispanic 
or Latino, `r comma(sum(az_dc_ndg$nhl_white))` people considered themselves White, 
`r comma(sum(az_dc_ndg$nhl_black))` people considered themselves Black or African-American, 
`r comma(sum(az_dc_ndg$nhl_ai_an))` people considered themselves American Indian or Alaskan 
Native, `r comma(sum(az_dc_ndg$nhl_asian))` people considered themselves Asian, 
`r comma(sum(az_dc_ndg$nhl_nhi_pi))` people considered themselves Native Hawaiian or Pacific 
Islander, and `r comma(sum(az_dc_ndg$pct_nhl_ao))`people considered themselves some other race 
or two or three or more races.

In those zip codes without COVID-19 data which maintained polling 
locations (n = `r nrow(az_dc_ndm)`),`r comma(sum(az_dc_ndm$total_hl))`people considered themselves 
Hispanic or Latino, `r comma(sum(az_dc_ndm$nhl_white))` people considered themselves White, 
`r comma(sum(az_dc_ndm$nhl_black))` people considered themselves Black or African-American, 
`r comma(sum(az_dc_ndm$nhl_ai_an))` people considered themselves American Indian or Alaskan 
Native, `r comma(sum(az_dc_ndm$nhl_asian))` people considered themselves Asian, 
`r comma(sum(az_dc_ndm$nhl_nhi_pi))` people considered themselves Native Hawaiian or Pacific 
Islander, and `r comma(sum(az_dc_ndm$pct_nhl_ao))` people considered themselves some other race 
or two or three or more races.

There were no zip codes without COVID-19 data which lost polling locations
between the 2016 and 2020 PPE.

*Maricopa County*

In those zip codes with complete COVID-19 data which gained polling locations 
(n = `r nrow(az_dcg_maricopa)`), `r comma(sum(az_dcg_maricopa$total_hl))`people considered 
themselves Hispanic or Latino, `r comma(sum(az_dcg_maricopa$nhl_white))` people considered 
themselves White, `r comma(sum(az_dcg_maricopa$nhl_black))` people considered themselves 
Black or African-American, `r comma(sum(az_dcg_maricopa$nhl_ai_an))` people considered 
themselves American Indian or Alaskan Native, `r comma(sum(az_dcg_maricopa$nhl_asian))` people 
considered themselves Asian, `r comma(sum(az_dcg_maricopa$nhl_nhi_pi))` people considered 
themselves Native Hawaiian or Pacific Islander, and `r comma(sum(az_dcg_maricopa$pct_nhl_ao))`
people considered themselves some other race or two or three or more races.

In those zip codes with complete COVID-19 data which maintained polling locations 
(n = `r nrow(az_dcm_maricopa)`),`r comma(sum(az_dcm_maricopa$total_hl))`people considered 
themselves Hispanic or Latino, `r comma(sum(az_dcm_maricopa$nhl_white))` people considered 
themselves White, `r comma(sum(az_dcm_maricopa$nhl_black))` people considered themselves 
Black or African-American, `r comma(sum(az_dcm_maricopa$nhl_ai_an))` people considered 
themselves American Indian or Alaskan Native, `r comma(sum(az_dcm_maricopa$nhl_asian))` people 
considered themselves Asian, `r comma(sum(az_dcm_maricopa$nhl_nhi_pi))` people considered 
themselves Native Hawaiian or Pacific Islander, and `r comma(sum(az_dcm_maricopa$pct_nhl_ao))`
people considered themselves some other race or two or three or more races.

In those zip codes with complete COVID-19 data which lost polling locations 
(n = `r nrow(az_dclost_maricopa)`),`r comma(sum(az_dclost_maricopa$total_hl))`people 
considered themselves Hispanic or Latino, `r comma(sum(az_dclost_maricopa$nhl_white))` 
people considered themselves White, `r comma(sum(az_dclost_maricopa$nhl_black))` people 
considered themselves Black or African-American, `r comma(sum(az_dclost_maricopa$nhl_ai_an))` 
people considered themselves American Indian or Alaskan Native, 
`r comma(sum(az_dclost_maricopa$nhl_asian))` people considered themselves Asian, 
`r comma(sum(az_dclost_maricopa$nhl_nhi_pi))` people considered themselves Native Hawaiian 
or Pacific Islander, and `r comma(sum(az_dclost_maricopa$pct_nhl_ao))`people considered 
themselves some other race or two or three or more races.

In those zip codes without complete COVID-19 data which gained polling locations 
(n = `r nrow(az_dcg_ndmaricopa)`), `r comma(sum(az_dcg_ndmaricopa$total_hl))`people considered 
themselves Hispanic or Latino, `r comma(sum(az_dcg_ndmaricopa$nhl_white))` people considered 
themselves White, `r comma(sum(az_dcg_ndmaricopa$nhl_black))` people considered themselves 
Black or African-American, `r comma(sum(az_dcg_ndmaricopa$nhl_ai_an))` people considered 
themselves American Indian or Alaskan Native, `r comma(sum(az_dcg_ndmaricopa$nhl_asian))` people 
considered themselves Asian, `r comma(sum(az_dcg_ndmaricopa$nhl_nhi_pi))` people considered 
themselves Native Hawaiian or Pacific Islander, and `r comma(sum(az_dcg_ndmaricopa$pct_nhl_ao))`
people considered themselves some other race or two or three or more races.

In those zip codes without complete COVID-19 data which maintained polling locations 
(n = `r nrow(az_dcm_ndmaricopa)`),`r comma(sum(az_dcm_ndmaricopa$total_hl))`people considered 
themselves Hispanic or Latino, `r comma(sum(az_dcm_ndmaricopa$nhl_white))` people considered 
themselves White, `r comma(sum(az_dcm_ndmaricopa$nhl_black))` people considered themselves 
Black or African-American, `r comma(sum(az_dcm_ndmaricopa$nhl_ai_an))` people considered 
themselves American Indian or Alaskan Native, `r comma(sum(az_dcm_ndmaricopa$nhl_asian))` people 
considered themselves Asian, `r comma(sum(az_dcm_ndmaricopa$nhl_nhi_pi))` people considered 
themselves Native Hawaiian or Pacific Islander, and `r comma(sum(az_dcm_ndmaricopa$pct_nhl_ao))`
people considered themselves some other race or two or three or more races.

There were no zip codes without COVID-19 data in Maricopa which lost polling 
locations between the 2016 and 2020 PPE.

############################################################

### South Carolina

In those zip codes with complete COVID-19 data which gained polling locations 
(n = `r nrow(sc_dc_g)`),`r comma(sum(sc_dc_g$total_hl))`people considered themselves Hispanic 
or Latino, `r comma(sum(sc_dc_g$nhl_white))` people considered themselves White, 
`r comma(sum(sc_dc_g$nhl_black))` people considered themselves Black or African-American, 
`r comma(sum(sc_dc_g$nhl_ai_an))` people considered themselves American Indian or Alaskan 
Native, `r comma(sum(sc_dc_g$nhl_asian))` people considered themselves Asian, 
`r comma(sum(sc_dc_g$nhl_nhi_pi))` people considered themselves Native Hawaiian or Pacific 
Islander, and `r comma(sum(sc_dc_g$pct_nhl_ao))`people considered themselves some other race 
or two or three or more races.

In those zip codes with complete COVID-19 data which maintained polling 
locations (n = `r nrow(sc_dc_m)`),`r comma(sum(sc_dc_m$total_hl))`people considered themselves 
Hispanic or Latino, `r comma(sum(sc_dc_m$nhl_white))` people considered themselves White, 
`r comma(sum(sc_dc_m$nhl_black))` people considered themselves Black or African-American, 
`r comma(sum(sc_dc_m$nhl_ai_an))` people considered themselves American Indian or Alaskan 
Native, `r comma(sum(sc_dc_m$nhl_asian))` people considered themselves Asian, 
`r comma(sum(sc_dc_m$nhl_nhi_pi))` people considered themselves Native Hawaiian or Pacific 
Islander, and `r comma(sum(sc_dc_m$pct_nhl_ao))`people considered themselves some other race 
or two or three or more races.

In those zip codes with complete COVID-19 data which lost polling locations 
(n = `r nrow(sc_dc_lost)`), `r comma(sum(sc_dc_lost$total_hl))`people considered themselves 
Hispanic or Latino, `r comma(sum(sc_dc_lost$nhl_white))` people considered themselves White, 
`r comma(sum(sc_dc_lost$nhl_black))` people considered themselves Black or African-American, 
`r comma(sum(sc_dc_lost$nhl_ai_an))` people considered themselves American Indian or Alaskan 
Native, `r comma(sum(sc_dc_lost$nhl_asian))` people considered themselves Asian, 
`r comma(sum(sc_dc_lost$nhl_nhi_pi))` people considered themselves Native Hawaiian or Pacific 
Islander, and `r comma(sum(sc_dc_lost$pct_nhl_ao))`people considered themselves some other 
race or two or three or more races.

*Richland County*

In those zip codes with complete COVID-19 data which gained polling locations 
(n = `r nrow(sc_dcg_richland)`), `r comma(sum(sc_dcg_richland$total_hl))`people considered 
themselves Hispanic or Latino, `r comma(sum(sc_dcg_richland$nhl_white))` people considered 
themselves White, `r comma(sum(sc_dcg_richland$nhl_black))` people considered themselves 
Black or African-American, `r comma(sum(sc_dcg_richland$nhl_ai_an))` people considered 
themselves American Indian or Alaskan Native, `r comma(sum(sc_dcg_richland$nhl_asian))` people 
considered themselves Asian, `r comma(sum(sc_dcg_richland$nhl_nhi_pi))` people considered 
themselves Native Hawaiian or Pacific Islander, and `r comma(sum(sc_dcg_richland$pct_nhl_ao))`
people considered themselves some other race or two or three or more races.

In those zip codes with complete COVID-19 data which maintained polling locations 
(n = `r nrow(sc_dcm_richland)`),`r comma(sum(sc_dcm_richland$total_hl))`people considered 
themselves Hispanic or Latino, `r comma(sum(sc_dcm_richland$nhl_white))` people considered 
themselves White, `r comma(sum(sc_dcm_richland$nhl_black))` people considered themselves 
Black or African-American, `r comma(sum(sc_dcm_richland$nhl_ai_an))` people considered 
themselves American Indian or Alaskan Native, `r comma(sum(sc_dcm_richland$nhl_asian))` people 
considered themselves Asian, `r comma(sum(sc_dcm_richland$nhl_nhi_pi))` people considered 
themselves Native Hawaiian or Pacific Islander, and `r comma(sum(sc_dcm_richland$pct_nhl_ao))`
people considered themselves some other race or two or three or more races.

In those zip codes with complete COVID-19 data which lost polling locations 
(n = `r nrow(sc_dclost_richland)`),`r comma(sum(sc_dclost_richland$total_hl))`people 
considered themselves Hispanic or Latino, `r comma(sum(sc_dclost_richland$nhl_white))` 
people considered themselves White, `r comma(sum(sc_dclost_richland$nhl_black))` people 
considered themselves Black or African-American, `r comma(sum(sc_dclost_richland$nhl_ai_an))` 
people considered themselves American Indian or Alaskan Native, 
`r comma(sum(sc_dclost_richland$nhl_asian))` people considered themselves Asian, 
`r comma(sum(sc_dclost_richland$nhl_nhi_pi))` people considered themselves Native Hawaiian 
or Pacific Islander, and `r comma(sum(sc_dclost_richland$pct_nhl_ao))`people considered 
themselves some other race or two or three or more races.

## Methodology

This report examines polling location data from `r nrow(ncountiesaz)`
or `r round_half_up(sum((nrow(ncountiesaz)/15)*100),0)`% of the counties in 
Arizona and `r nrow(ncountiessc)` or `r round_half_up(sum((nrow(ncountiessc)/46)*100),0)`% 
of the counties in South Carolina. Pima County, AZ has not been included since 
the data come from the Voter Information Project and Pima County does not 
participate in its activities. Data were reviewed for uniqueness, completeness, 
and accuracy of information on zip codes, county, and number of polling locations. 
Polling locations are known to change just before an election so this data 
should not be viewed as a definitive list of the actual polling locations used 
on election day.

Demographic data were obtained from the 2014-2018 American Community Survey's 
B03002 table and groups were categorized as either Hispanic or Latino, and
not Hispanic or Latino by specific group: White alone, Black or African American 
Alone, American Indian and Alaska Native alone, Asian alone,Native Hawaiian and 
Other Pacific Islander alone, and All other races alone (Including Some other 
race alone and Two or more races alone).

Zip code and county data were obtained from unitedstateszipcodes.org. 
While many zip codes map directly to specific counties, the relationship between 
these two organizational systems is not always 1:1 and is subject to change.

Data on COVID-19 cases in Arizona was obtained from the Arizona Department 
of Health Services while data on cases in South Carolina was obtained
though the New York Times covid-19 repository on GitHub. Data from these sources
was combined with demographic and zip code data. Data on COVID-19 cases was only 
kept for those zip codes for which we also had polling location data and in 
Arizona, was categorized to denote where data were missing or suppressed pending 
tribal approval as noted in the original data. Data are current as of as 
of `r format(Sys.time(), '%d %B, %Y')`.

## Acknowledgements {-} 

We are grateful to Sarah Berlin at Democracy Works for collecting data from the 
Voter Information Project.